---
title: "In vitro Beta Diversity T48"
author: "apih"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    toc_depth: 4
    collapsed: false
    code_folding: hide
    number_sections: false
knit: (function(inputFile, encoding) { 
    rmarkdown::render(
        inputFile, encoding = encoding,
        output_dir = file.path(dirname(inputFile),"output"),
        output_file = paste0("PRIMA06_", Sys.Date(), "_beta_diversity_groups.html"))
        })
params:
    input: "R_objects/Phyloseq_harsh_invitro.Rdata"
    group_var: "Treatment"
    subject_var: "Tube"
    time_var: "Time_point"
---

# INFO

This template is build to use the output from **GMH_import** as input. It will take you through a basic beta diversity analysis for the project. Beta diversity is the difference between samples, this means that we compare samples as a whole. The analysis generally consists of three steps:

1.  Calculation of distance/dissimilarity between samples.

2.  Visualization of the data (ordination).

3.  Statistical analysis (PERMANOVA).


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(GMHmicrobiome)
library(ggpubr)
library(kableExtra)
library(phyloseq)
library(cowplot)
library(ggExtra)
library(vegan)
library(ecole)


# save parameters
saveRDS(params, "R_objects/params_betadiv_invitro.RDS")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


# CALCULATE DISTANCES

## PREPARE DATA

First step is to create a clean phyloseq object (remove samples missing relevant data and/or subset by a relevant variable).

Beta diversity are affected by all samples included, so if some samples are removed, this should preferably be done before the beta diversity is calculated. \> qualitative metrics are sensitive to sequencing depth, so rarefaction should be done for those metrics as well.

```{r prepare-data, eval=FALSE, echo=TRUE}

# Load data
params <- readRDS("R_objects/params_betadiv_invitro.RDS")
load(params$input)

# Create subset
phy.clean <- subset_samples(phy, Time_point == "T48")
phy.clean <- subset_samples(phy.clean, Treatment %in% c("Saline", "Imolope_low", "Imolope_high"))

# Clean data
phy.clean <- prune_samples(sample_names(phy.clean)[!is.na(sample_data(phy.clean)[,params$group_var])], phy.clean)

# Remove empty taxa
phy.clean <- prune_taxa(taxa_sums(phy.clean) > 0, phy.clean)

# Recalculate midpoint root
phy_tree(phy.clean) <- phangorn::midpoint(phy_tree(phy.clean))

# Perform multiple rarefactions
phy.rare <- multiple_rarefy(phy.clean)

# Remove empty taxa
phy.rare <- prune_taxa(taxa_sums(phy.rare) > 0, phy.rare)

# Recalculate midpoint root
phy_tree(phy.rare) <- phangorn::midpoint(phy_tree(phy.rare))

# Save object
save(phy.rare, phy.clean, file = "R_objects/Phyloseq_betadiv_T48.Rdata")

# clear the environment and release memory
rm(list = ls(all.names = TRUE)) #will clear all objects includes hidden objects.
invisible(gc()) #free up memory and report the memory usage.
```

## CALCULATE

### AITCHISON

Aitchison distance (1986) and robust Aitchison distance [(Martino et al. 2019)](https://journals.asm.org/doi/10.1128/mSystems.00016-19) are metrics that deal with compositional data. The Aitchison distance is a dissimilarity measure calculated as the Euclidean distance between observations (samples) after performing a centered log ratio (“clr”) transformation. Aitchison distance has been said to outperform Jensen-Shannon divergence and Bray-Curtis dissimilarity, due to a better stability to subsetting and aggregation, and it being a proper distance (Aitchison et al., 2000).

```{r calc-ait, eval=FALSE, echo=TRUE}

# load
params <- readRDS("R_objects/params_betadiv_invitro.RDS")
load("R_objects/Phyloseq_betadiv_T48.Rdata")

# Calculate Bray-Curtis dissimilarities
ait.dist <- vegan::vegdist(otu_table(phy.clean), method = "robust.aitchison")

# Calculate PCoA data
ait.pcoa <- ordinate(phy.clean, method = "PCoA",distance = ait.dist)
ait.nmds <- metaMDS(ait.dist, k = 5, trymax = 1000)

# Save distance objects
save(ait.dist, ait.nmds, ait.pcoa, phy.clean, file = "R_objects/bdiv_ait_T48.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


### JACCARD

The Jaccard similarity measures the similarity between two sets of data to see which members are shared and distinct. The Jaccard similarity is calculated by dividing the number of observations in both sets by the number of observations in either set. In other words, the Jaccard similarity can be computed as the size of the intersection divided by the size of the union of two sets. This can be written in set notation using intersection $(A \cap B)$ and unions $(A \cup B)$ of two sets: $$J(A,B) = \frac{|A \cap B|}{|A \cup B|}$$ where $|A \cap B|$ gives the number of members shared between both sets and $|A \cup B|$ gives the total number of members in both sets (shared and un-shared). The Jaccard Similarity will be 0 if the two sets don't share any values and 1 if the two sets are identical.

Additionally, this function can be used to find the dissimilarity between two sets by calculating:
$$d(A,B) = 1 - J(A,B)$$
> We will be calculating binary Jaccard dissimilarity

```{r calc-jac, eval=FALSE, echo=TRUE}

# load
params <- readRDS("R_objects/params_betadiv_invitro.RDS")
load("R_objects/Phyloseq_betadiv_T48.Rdata")

# Calculate Jaccard binary dissimilarities
jac.dist <- distance(phy.rare, method = "jaccard", binary = TRUE)

# Calculate PCoA data
jac.pcoa <- ordinate(phy.rare, method = "PCoA",distance = jac.dist)
jac.nmds <- metaMDS(jac.dist, k = 5, trymax = 1000)

# Save distance objects
save(jac.dist, jac.nmds, jac.pcoa, phy.rare, file = "R_objects/bdiv_jac_T48.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


# Analysis 1: Aitchison 

## VISUALIZATION

### LOAD DATA

```{r load-betadiv-a, eval=TRUE, echo=TRUE}

params <- readRDS("R_objects/params_betadiv_invitro.RDS")

# Choose metric
METRIC <- "ait"

# Choose variable 
VAR <- "Treatment"

# Load data
load(params$input)

if (METRIC == "jac") {
  load("R_objects/bdiv_jac_T48.RData")
  dist.used <- bray.dist
  nmds.used <- bray.nmds
  pcoa.used <- bray.pcoa
  phy.used <- phy.rare
  rm(jac.dist, jac.nmds, jac.pcoa, phy.rare)
} else if (METRIC == "ait"){
  load("R_objects/bdiv_ait_T48.RData")
  dist.used <- ait.dist
  nmds.used <- ait.nmds
  pcoa.used <- ait.pcoa
  phy.used <- phy.clean
  rm(ait.dist, ait.nmds, ait.pcoa, phy.clean)
}

# Extract metadata from phyloseq
mdat <- data.frame(sample_data(phy.used))

# If a variable consist of numbers, but represent distinct groups remember to make it into a factor
mdat[,VAR] <- as.factor(mdat[,VAR])
```

### CLEAN ORDINATION {.tabset .tabset-fade .tabset-pills}

#### EIGENVALUES

```{r plot-eigenvalues-a, eval=TRUE, echo=TRUE}

# Extract eigen values
eigen <- pcoa.used$values
eigen$Axis <- as.numeric(row.names(eigen))

head(eigen)

# Create plots for both distance indeces
p.eigen <- ggplot(eigen[1:10,], aes(x = as.factor(Axis), y = 100*Relative_eig)) + 
  geom_col(aes(fill = as.factor(Axis))) +
  geom_point(aes(x = Axis, y = 100*Cumul_eig)) +
  geom_line(aes(x = Axis, y = 100*Cumul_eig)) +
  ylab("Variance explained (%)") +
  xlab("Axis") +
  theme_pubr(legend = "none") + ggsci::scale_fill_jco()
p.eigen 

```

#### PLOT

```{r ordination-find-axis-a, eval=TRUE}

# Create plots of eigenvalues for PCoA plots
pcoa.tab <- plot_ordination(phy, pcoa.used,axes = 1:5,justDF = TRUE)
nmds.tab <- plot_ordination(phy, nmds.used,axes = 1:5,justDF = TRUE)

# Reformat tables to create one common table
colnames(nmds.tab)[1:5] <- c("Axis.1","Axis.2","Axis.3","Axis.4","Axis.5")

nmds.tab$ordination <- "nmds"
pcoa.tab$ordination <- "pcoa"

ord.tab <- rbind(nmds.tab,pcoa.tab)
ord.tab[,VAR] <- as.factor(ord.tab[,VAR])

# Melt axis to be in one variable
axis.tab <- pivot_longer(data = ord.tab, cols = c("Axis.1","Axis.2","Axis.3","Axis.4","Axis.5"), names_to = "Axis", values_to = "position")

# Plot positions on axes
ggplot(axis.tab, aes_string(x = "ordination", y = "position", fill = VAR)) +
  geom_boxplot() +
  facet_grid(Axis~.) +
  coord_flip() + 
  theme_pubr(legend = "bottom") + ggsci::scale_fill_jco()

```


```{r plot-ordination-a, eval=TRUE}

#Sort group variable
str(ord.tab$Treatment)
ord.tab$Treatment <- factor(ord.tab$Treatment, levels = c("Saline", "Imolope_low", "Imolope_high"), ordered = TRUE)

my_cols <- c("#CC3333","#0066CC", "#88419d")

# Create PCoA plot 
plot.pcoa1 <- ggplot(ord.tab[ord.tab$ordination == "pcoa",], aes_string(x = "Axis.1", y = "Axis.2", color = VAR, group = VAR)) +
  geom_point() + 
  labs(color=VAR) +
  theme_pubr(legend = "bottom")+
  stat_ellipse(aes_string(x = "Axis.1", y = "Axis.2", color = VAR, fill = VAR),geom = "polygon", alpha = 0.1, level = 0.9) +
  scale_color_manual(values = my_cols)+
  scale_fill_manual(values = my_cols)+
  border(color = "black",size = 0.5) +
  labs(caption = "PCoA")+
  #xlab("Axis 1 (12%)")+
  #ylab("Axis 2 (10%)")+ 
  theme(
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size=11),
    axis.text.y = element_text(size=10),
    axis.title.y = element_text(size = 11), 
    legend.text = element_text(size = 10), 
    legend.title = element_blank())


# Add marginal box plot 
plot.pcoa.box <- ggExtra::ggMarginal(p = plot.pcoa1, type = 'boxplot', size = 10, groupFill = TRUE)

plot.pcoa.box <- annotate_figure(
  plot.pcoa.box,
  top = text_grob("Aitchison distance 48h (ASVs)", face = "bold", size = 12),
)

plot.pcoa.box


filename <- paste0("plots/In_vitro/aitchison_t48_asv.png")
plot.pcoa.box
suppressMessages(ggsave(filename = filename, width = 1800, height = 1400, units = "px", dpi = "print", device = "png", ))

```


## STATISTICAL ANALYSIS

### PERMANOVA

#### DISPERTION

First we will test the beta diversity dispertion to determine whether any differences in dispertion might cause any PERMANOVA differences.

```{r betadisp-a, eval=TRUE}

# Calculate betadispertion
bdisp <- betadisper(dist.used, mdat[,VAR])

# Run statical test
anova(bdisp)

# Run posthoc test if significant and more than two groups
TukeyHSD(bdisp)
plot(TukeyHSD(bdisp))

# Plot dispertion
boxplot(bdisp)

```

#### RUN TEST

```{r permanova-a, eval=TRUE}

# set seeds
set.seed(9)

# Pairwise test
permanova_pairwise(x = dist.used, grp = mdat$Treatment, padj = "fdr")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

# Analysis 2: Jaccard  

## VISUALIZATION

### LOAD DATA

```{r load-betadiv-b, eval=TRUE, echo=TRUE}

params <- readRDS("R_objects/params_betadiv_invitro.RDS")

# Choose metric
METRIC <- "jac"

# Choose variable 
VAR <- "Treatment"

# Load data
load(params$input)

if (METRIC == "jac") {
  load("R_objects/bdiv_jac_T48.RData")
  dist.used <- jac.dist
  nmds.used <- jac.nmds
  pcoa.used <- jac.pcoa
  phy.used <- phy.rare
  rm(jac.dist, jac.nmds, jac.pcoa, phy.rare)
} else if (METRIC == "ait"){
  load("R_objects/bdiv_ait_T48.RData")
  dist.used <- ait.dist
  nmds.used <- ait.nmds
  pcoa.used <- ait.pcoa
  phy.used <- phy.clean
  rm(ait.dist, ait.nmds, ait.pcoa, phy.clean)
}

# Extract metadata from phyloseq
mdat <- data.frame(sample_data(phy.used))

# If a variable consist of numbers, but represent distinct groups remember to make it into a factor
mdat[,VAR] <- as.factor(mdat[,VAR])
```

### CLEAN ORDINATION {.tabset .tabset-fade .tabset-pills}

#### EIGENVALUES

```{r plot-eigenvalues-b, eval=TRUE, echo=TRUE}

# Extract eigen values
eigen <- pcoa.used$values
eigen$Axis <- as.numeric(row.names(eigen))

head(eigen)

# Create plots for both distance indeces
p.eigen <- ggplot(eigen[1:10,], aes(x = as.factor(Axis), y = 100*Relative_eig)) + 
  geom_col(aes(fill = as.factor(Axis))) +
  geom_point(aes(x = Axis, y = 100*Cumul_eig)) +
  geom_line(aes(x = Axis, y = 100*Cumul_eig)) +
  ylab("Variance explained (%)") +
  xlab("Axis") +
  theme_pubr(legend = "none") + ggsci::scale_fill_jco()
p.eigen 

```

#### PLOT

```{r ordination-find-axis-b, eval=TRUE}
# Create plots of eigenvalues for PCoA plots
pcoa.tab <- plot_ordination(phy, pcoa.used,axes = 1:5,justDF = TRUE)
nmds.tab <- plot_ordination(phy, nmds.used,axes = 1:5,justDF = TRUE)

# Reformat tables to create one common table
colnames(nmds.tab)[1:5] <- c("Axis.1","Axis.2","Axis.3","Axis.4","Axis.5")

nmds.tab$ordination <- "nmds"
pcoa.tab$ordination <- "pcoa"

ord.tab <- rbind(nmds.tab,pcoa.tab)
ord.tab[,VAR] <- as.factor(ord.tab[,VAR])

# Melt axis to be in one variable
axis.tab <- pivot_longer(data = ord.tab, cols = c("Axis.1","Axis.2","Axis.3","Axis.4","Axis.5"), names_to = "Axis", values_to = "position")

# Plot positions on axes
ggplot(axis.tab, aes_string(x = "ordination", y = "position", fill = VAR)) +
  geom_boxplot() +
  facet_grid(Axis~.) +
  coord_flip() + 
  theme_pubr(legend = "bottom") + ggsci::scale_fill_jco()
```


```{r plot-ordination-b, eval=TRUE}

#Sort group variable
str(ord.tab$Treatment)
ord.tab$Treatment <- factor(ord.tab$Treatment, levels = c("Saline", "Imolope_low", "Imolope_high"), ordered = TRUE)

my_cols <- c("#CC3333","#0066CC", "#88419d")

# Create PCoA plot 
plot.pcoa1 <- ggplot(ord.tab[ord.tab$ordination == "pcoa",], aes_string(x = "Axis.1", y = "Axis.2", color = VAR, group = VAR)) +
  geom_point() + 
  labs(color=VAR) +
  theme_pubr(legend = "bottom")+
  stat_ellipse(aes_string(x = "Axis.1", y = "Axis.2", color = VAR, fill = VAR),geom = "polygon", alpha = 0.1, level = 0.9) +
  scale_color_manual(values = my_cols)+
  scale_fill_manual(values = my_cols)+
  border(color = "black",size = 0.5) +
  labs(caption = "PCoA")+
  #xlab("Axis 1 (14%)")+
  #ylab("Axis 2 (9%)")+ 
  theme(
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size=11),
    axis.text.y = element_text(size=10),
    axis.title.y = element_text(size = 11), 
    legend.text = element_text(size = 10), 
    legend.title = element_blank())
  
# Add marginal box plot 
plot.pcoa.box <- ggExtra::ggMarginal(p = plot.pcoa1, type = 'boxplot', size = 10, groupFill = TRUE)

plot.pcoa.box <- annotate_figure(
  plot.pcoa.box,
  top = text_grob("Jaccard 48h (ASVs)", face = "bold", size = 12),
)

plot.pcoa.box

filename <- paste0("plots/In_vitro/jaccard_t48_asv.png")
plot.pcoa.box
suppressMessages(ggsave(filename = filename, width = 1800, height = 1400, units = "px", dpi = "print", device = "png", ))

```


## STATISTICAL ANALYSIS

### PERMANOVA

#### DISPERTION

First we will test the beta diversity dispertion to determine whether any differences in dispertion might cause any PERMANOVA differences.

```{r betadisp-b, eval=TRUE}

# Calculate betadispertion
bdisp <- betadisper(dist.used, mdat[,VAR])

# Run statical test
anova(bdisp)

# Run posthoc test if significant and more than two groups
TukeyHSD(bdisp)
plot(TukeyHSD(bdisp))

# Plot dispertion
boxplot(bdisp)

```

#### RUN TEST

```{r permanova-b, eval=TRUE}

# set seeds
set.seed(9)

# Pairwise test
permanova_pairwise(x = dist.used, grp = mdat$Treatment, padj = "fdr")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```



## SESSION INFO

The analysis was run in the following environment:

```{r packages, eval=TRUE}
sessionInfo()
```
