---
title: "In vitro alpha diversity"
author: "apih"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    toc_depth: 4
    collapsed: false
    code_folding: hide
    number_sections: false
knit: (function(inputFile, encoding) { 
    rmarkdown::render(
        inputFile, encoding = encoding,
        output_dir = file.path(dirname(inputFile),"output"),
        output_file = paste0("PRIMA06_", Sys.Date(), "_test_variables.html"))
        })
---

# INFO {.tabset .tabset-fade .tabset-pills}

This Rmarkdown contains the commands necessary to perform alpha diversity analysis of the output from the [DF_GMH_PIPELINE](https://github.com/MSMortensen/DF_GMH_pipeline).

Alpha diversity, also called "within sample diversity" is calculated for each sample individually and is independent of all other samples. Alpha diversity is sensitive to sequencing depth, so rarefaction must be done first.
Here we calculate index:
 -    Observed richness (observed species)
 -    Shannon diversity index

# SETUP

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(GMHmicrobiome)
library(ggpubr)
library(kableExtra)
library(phyloseq)
library(rstatix)
library(vegan)

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


# DATA

To ensure the data is in the right format, this section will load, view, clean, and save the data for use throughout this template.

## Load

```{r load, eval=TRUE}

# Load phyloseq metadata
load("R_objects/Phyloseq_harsh_invitro.Rdata")

# Extract sample data from phyloseq object
dat <- as_tibble(sample_data(phy))


```

## Look

It is important to understand the structure and format of your data.

```{r look, eval=TRUE}

# Take a glimpse
glimpse(dat)

View(dat)

# Make a explorative summary 
skimr::skim(dat)

```

## Clean

If there are data that is poorly structured or wrongly formatted, it will be sorted here.

This is the general principles to follow:

-   Remove variables

    -   With one unique value (e.g. sample_type, is_neg)

    -   With duplicated information (e.g. keep one of sampleID and ID)

    -   With irrelevant data (e.g. barcode)

-   Change format

    -   Numeric values as numbers `<dbl>` or integers `<int>.`

    -   Unique strings as characters `<chr>`

    -   Categorical data as factors `<fct>` (unordered, e.g. treatment groups)

    -   Ordered categorical data as ordered `<ord>` (e.g. size: "small" \< "medium" \< "large")

```{r clean, eval=TRUE}

# Remove
out.var <- c("type","is.neg", "Sample","ASVs","Seq_run", "Project")

dat <- dat %>% select(-one_of(out.var))

# I will add a column with tube identifier that I forgot to add in the metadata in the first place 
metadat <- read_csv("input/lop_invitro_meta.csv")

# remove mock and negative control 
metadat <- metadat[!is.na(metadat[,"Tube"]),]

# Sort both data frames 
metadat <- metadat[order(metadat$Sample), ]
dat <- dat[order(dat$SampleID), ]

# Add the extra variable column
dat$Tube <- metadat$Tube

# Subset data 
dat <- subset(dat, Treatment %in% c("Saline", "Imolope_low", "Imolope_high"))

# For ordered factors I suggest to do them individually
dat <- dat %>% mutate(Time_point = factor(Time_point, levels = c("T0","T24","T48", "T72"),ordered = TRUE))
dat <- dat %>% mutate(Treatment = factor(Treatment, levels = c("Saline", "Imolope_low", "Imolope_high"),ordered = TRUE))
dat <- dat %>% mutate(Date = factor(Date, levels = c("04_03_25","05_03_25","06_03_25", "07_03_25"),ordered = TRUE))
dat <- dat %>% mutate_at("Tube_ID", factor)
dat <- dat %>% mutate_at("Tube", factor)

# Look at cleaned data
skimr::skim(dat)

```

## Save

The cleaned data is here saved in one file that can be loaded when necessary in this template. If you need subsets for the analysis I suggest to create and save those here as well

```{r save, eval=TRUE}
# Save cleaned data
save(dat, file = "R_objects/alpha_data_invitro.RData")


# clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```

# OBSERVED

```{r prepare_observed, eval=TRUE}
# load data 
load("R_objects/alpha_data_invitro.RData")

# Set names of variables
PREDICTOR <- "Treatment"
OUTCOME <- "Observed"

dat.clean <- dat


dat.t24 <- dat.clean[dat.clean$Time_point %in% c('T24'),]

dat.t48 <- dat.clean[dat.clean$Time_point %in% c('T48'),] 

dat.t72 <- dat.clean[dat.clean$Time_point %in% c('T72'),] 



```

#### Visualize and summarize
```{r visualize_observed}

# Summary samples in groups
dat.clean %>% 
  group_by(!!sym(PREDICTOR)) %>% 
  get_summary_stats(!!sym(OUTCOME), type = "common")


# Set colors 
my_cols <- c("#CC3333","#0066CC", "#88419d")


bxp_t24 <- dat.t24 %>%
  ggboxplot(x = PREDICTOR,
           y = OUTCOME,
           fill = PREDICTOR) + 
  geom_point(size = 1, color = "black")+
  ylim(60,110) +
  ylab("Observed OTUs")+
  ggtitle("24 hours")+
  theme_classic()+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))+
  scale_fill_manual(values=my_cols, labels = c("Saline vehicle", "Saline + low imolope", "Saline + high imolope"))

bxp_t48 <- dat.t48 %>%
  ggboxplot(x = PREDICTOR,
           y = OUTCOME,
           fill = PREDICTOR) + 
  geom_point(size = 1, color = "black")+
  ylim(60,110) +
  ggtitle("48 hours")+
  ylab(NULL)+
  theme_classic()+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.line.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())+
   stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))+
  scale_fill_manual(values=my_cols, labels = c("Saline vehicle", "Saline + low imolope", "Saline + high imolope"))


bxp_t72 <- dat.t72 %>%
  ggboxplot(x = PREDICTOR,
           y = OUTCOME,
           fill = PREDICTOR) + 
  geom_point(size = 1, color = "black")+
  ylim(60,110) +
  ggtitle("72 hours")+
  ylab(NULL)+
  theme_classic()+
  theme(legend.position="right")+ 
  theme(axis.title.x=element_blank(), 
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.line.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))+
  scale_fill_manual(values=my_cols, labels = c("Saline vehicle", "Saline + low imolope", "Saline + high imolope"))




```


#### Check normality and homogneity of variance assumptions

QQ plot and Shapiro-Wilk test of normality are used to analyze the model residuals.
The homogeneity of variance assumption of the between-subject factor can be checked using the Levene's test. The test is performed at each level of within-subject variable.

```{r assumptions_observed}

# Create a QQ plot of residuals

dat.t24 %>%
  ggqqplot(OUTCOME)

dat.t48 %>%
  ggqqplot(OUTCOME)

dat.t72 %>%
  ggqqplot(OUTCOME)


# Compute Shapiro-Wilk test of normality

dat.t24 %>%
  shapiro_test(!!sym(OUTCOME))

dat.t48 %>%
  shapiro_test(!!sym(OUTCOME))

dat.t72 %>%
  shapiro_test(!!sym(OUTCOME))


# Levene's test

dat.t24 %>%
  levene_test(as.formula(paste(OUTCOME,PREDICTOR,sep = " ~ ")))

dat.t48 %>%
  levene_test(as.formula(paste(OUTCOME,PREDICTOR,sep = " ~ ")))

dat.t72 %>%
  levene_test(as.formula(paste(OUTCOME,PREDICTOR,sep = " ~ ")))



# All assumptions are met 

```


#### Statistical tests 

```{r stat_tests_observed}
# ANOVA
# Two-way anova for repeated measures 
two.way <- anova_test(data = dat, 
                      dv = !!sym(OUTCOME), 
                      wid = !!sym("Tube"), 
                      between = !!sym(PREDICTOR),
                      within = !!sym("Time_point"))
get_anova_table(two.way) # Significant difference for Time point 


# Post hoc
stat.test.t24 <- dat.t24 %>% 
  t_test(Observed~Treatment, paired = FALSE, p.adjust.method = "fdr") %>% 
  add_significance("p.adj") %>%
  add_xy_position(x = "Treatment", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE) # not significant 


stat.test.t48 <- dat.t48 %>% 
  t_test(Observed~Treatment, paired = FALSE, p.adjust.method = "fdr") %>% 
  add_significance("p.adj") %>%
  add_xy_position(x = "Treatment", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE) # Significance 


stat.test.t72 <- dat.t72 %>% 
  t_test(Observed~Treatment, paired = FALSE, p.adjust.method = "fdr") %>% 
  add_significance("p.adj") %>%
  add_xy_position(x = "Treatment", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE) # not significant 



```


#### Report

```{r report_observed}
# Visualization: boxplots with p-values
My_Theme = theme(plot.title = element_text(size = 10)) # For setting the text sizes in the plot 


plot_t24_f <- bxp_t24 + My_Theme + stat_pvalue_manual(stat.test.t24, label = "p.adj.signif", tip.length = 0.01, size = 3.5, hide.ns = TRUE)


plot_t48_f <- bxp_t48 + My_Theme + stat_pvalue_manual(stat.test.t48, label = "p.adj.signif", tip.length = 0.01, size = 3.5, hide.ns = TRUE, y.position = c(104,108)) 


plot_t72_f <- bxp_t72 + My_Theme + stat_pvalue_manual(stat.test.t72, label = "p.adj.signif", tip.length = 0.01, size = 3.5, hide.ns = TRUE)


plot_final <- ggarrange(plot_t24_f, plot_t48_f, plot_t72_f, ncol=3, nrow=1, widths = c(1, 0.8, 1.7))

plot_final <- annotate_figure(plot_final, top = text_grob("Observed OTUs", color = "black", face = "bold", hjust = 2.75))

filename <- paste0("plots/In_vitro/Observed.png")
plot_final
suppressMessages(ggsave(filename = filename, width = 2400, height = 1400, units = "px", dpi = "print", device = "png", ))

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())


```

# SHANNON

```{r prepare_shannon, eval=TRUE}
# load data 
load("R_objects/alpha_data_invitro.RData")

# Set names of variables
PREDICTOR <- "Treatment"
OUTCOME <- "Shannon"

dat.clean <- dat


dat.t24 <- dat.clean[dat.clean$Time_point %in% c('T24'),]

dat.t48 <- dat.clean[dat.clean$Time_point %in% c('T48'),]

dat.t72 <- dat.clean[dat.clean$Time_point %in% c('T72'),] 


```

#### Visualize and summarize
```{r visualize_shannon}

# Summary samples in groups
dat.clean %>% 
  group_by(!!sym(PREDICTOR)) %>% 
  get_summary_stats(!!sym(OUTCOME), type = "common")


# Set colors 
my_cols <- c("#CC3333","#0066CC", "#88419d")

bxp_t24 <- dat.t24 %>%
  ggboxplot(x = PREDICTOR,
           y = OUTCOME,
           fill = PREDICTOR) + 
  geom_point(size = 1, color = "black")+
  ylim(2, 2.72) +
  ylab("Shannon index")+
  ggtitle("24 hours")+
  theme_classic()+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))+
  scale_fill_manual(values=my_cols, labels = c("Saline vehicle", "Saline + low imolope", "Saline + high imolope"))

bxp_t48 <- dat.t48 %>%
  ggboxplot(x = PREDICTOR,
           y = OUTCOME,
           fill = PREDICTOR) + 
  geom_point(size = 1, color = "black")+
  ylim(2, 2.72) +
  ggtitle("48 hours")+
  ylab(NULL)+
  theme_classic()+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.line.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())+
   stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))+
  scale_fill_manual(values=my_cols, labels = c("Saline vehicle", "Saline + low imolope", "Saline + high imolope"))


bxp_t72 <- dat.t72 %>%
  ggboxplot(x = PREDICTOR,
           y = OUTCOME,
           fill = PREDICTOR) + 
  geom_point(size = 1, color = "black")+
  ylim(2, 2.72) +
  ggtitle("72 hours")+
  ylab(NULL)+
  theme_classic()+
  theme(legend.position="right")+ 
  theme(axis.title.x=element_blank(), 
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.line.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))+
  scale_fill_manual(values=my_cols, labels = c("Saline vehicle", "Saline + low imolope", "Saline + high imolope"))





```


#### Check normality and homogneity of variance assumptions

QQ plot and Shapiro-Wilk test of normality are used to analyze the model residuals.
The homogeneity of variance assumption of the between-subject factor can be checked using the Levene's test. The test is performed at each level of within-subject variable.

```{r assumptions_shannon}

# Create a QQ plot of residuals
dat.t24 %>%
  ggqqplot(OUTCOME)

dat.t48 %>%
  ggqqplot(OUTCOME)

dat.t72 %>%
  ggqqplot(OUTCOME)


# Compute Shapiro-Wilk test of normality
dat.t24 %>%
  shapiro_test(!!sym(OUTCOME))

dat.t48 %>%
  shapiro_test(!!sym(OUTCOME))

dat.t72 %>%
  shapiro_test(!!sym(OUTCOME))



# Levene's test
dat.t24 %>%
  levene_test(as.formula(paste(OUTCOME,PREDICTOR,sep = " ~ ")))

dat.t48 %>%
  levene_test(as.formula(paste(OUTCOME,PREDICTOR,sep = " ~ ")))

dat.t72 %>%
  levene_test(as.formula(paste(OUTCOME,PREDICTOR,sep = " ~ ")))



# All assumptions are met 

```


#### Statistical tests 

```{r stat_tests_shannon}
# ANOVA
# Two-way anova for repeated measures 
two.way <- anova_test(data = dat, 
                      dv = !!sym(OUTCOME), 
                      wid = !!sym("Tube"), 
                      between = !!sym(PREDICTOR),
                      within = !!sym("Time_point"))
get_anova_table(two.way) # Significant difference for Time point, Treatment and interaction  


# Post hoc
stat.test.t24 <- dat.t24 %>% 
  t_test(Shannon~Treatment, paired = FALSE, p.adjust.method = "fdr") %>% 
  add_significance("p.adj") %>%
  add_xy_position(x = "Treatment", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE) # not significant 

stat.test.t48 <- dat.t48 %>% 
  t_test(Shannon~Treatment, paired = FALSE, p.adjust.method = "fdr") %>% 
  add_significance("p.adj") %>%
  add_xy_position(x = "Treatment", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE) # not significant 

stat.test.t72 <- dat.t72 %>% 
  t_test(Shannon~Treatment, paired = FALSE, p.adjust.method = "fdr") %>% 
  add_significance("p.adj") %>%
  add_xy_position(x = "Treatment", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE) # significant 



```


#### Report

```{r report_shannon}
# Visualization: boxplots with p-values
My_Theme = theme(plot.title = element_text(size = 10)) # For setting the text sizes in the plot 


plot_t24_f <- bxp_t24 + My_Theme + stat_pvalue_manual(stat.test.t24, label = "p.adj.signif", tip.length = 0.01, size = 3.5, hide.ns = TRUE)


plot_t48_f <- bxp_t48 + My_Theme + stat_pvalue_manual(stat.test.t48, label = "p.adj.signif", tip.length = 0.01, size = 3.5, hide.ns = TRUE) 


plot_t72_f <- bxp_t72 + My_Theme + stat_pvalue_manual(stat.test.t72, label = "p.adj.signif", tip.length = 0.01, size = 3.5, hide.ns = TRUE, y.position = c(2.68, 2.72))


plot_final <- ggarrange(plot_t24_f, plot_t48_f, plot_t72_f, ncol=3, nrow=1, widths = c(1, 0.8, 1.7))

plot_final <- annotate_figure(plot_final, top = text_grob("Shannon diversity", color = "black", face = "bold", hjust = 2.41))

filename <- paste0("plots/In_vitro/Shannon.png")
plot_final
suppressMessages(ggsave(filename = filename, width = 2400, height = 1400, units = "px", dpi = "print", device = "png", ))


# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())


```

# INFO {.tabset .tabset-fade .tabset-pills}

## FINAL COMMENT

This completes the statistical analysis of alpha diversity 

## SESSION INFO

The analysis was run in the following environment:

```{r packages, eval=TRUE}
sessionInfo()
```
