---
title: "In vitro absolute abundance"
author: "APIH"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    toc_depth: 4
    collapsed: false
    code_folding: hide
    number_sections: false
knit: (function(inputFile, encoding) { 
    rmarkdown::render(
        inputFile, encoding = encoding,
        output_dir = file.path(dirname(inputFile),"output"),
        output_file = paste0("PRIMA08", Sys.Date(), "_differential_abundance.html"))
        })
params:
    input: "R_objects/Phyloseq_invitro.Rdata"
    group_var: "Treatment"
    subject_var: "Tube"
    time_var: "Time_point"
---


```{r setup, eval=TRUE, echo=TRUE, message=FALSE,warning = FALSE}
knitr::opts_chunk$set(echo = TRUE ,warning = FALSE, message = FALSE)

# Load libraries
library(tidyverse)
library(phyloseq)
library(decontam)
library(pals)
library(ggpubr)
library(vegan)
library(phangorn)
library(kableExtra)
library(DAtest)
library(rstatix)
library(ggplot2)

# save parameters
saveRDS(params, "R_objects/params_DA_invitro.RDS")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


## FAMILY TRAJECTORIES 
### Bacteroidaceae
#### Prepare dataset 
```{r Bacteroidaceae_prep, eval = TRUE}
# Load data 
dat <- read_csv("output/FamilyAbundance_invitro.csv")

# Select data 
dat <- dat[dat$Family %in% c('Bacteroidaceae'),] 
dat <- subset(dat, Treatment %in% c("Saline", "Imolope_low", "Imolope_high"))

# Add a column with tube identifier that I forgot to add in the metadata in the first place 
# Read metadata 
metadat <- read_csv("input/lop_invitro_meta.csv")
metadat <- subset(metadat, Treatment %in% c("Saline", "Imolope_low", "Imolope_high"))

# remove mock and negative control 
metadat <- metadat[!is.na(metadat[,"Tube"]),]

# Sort both data frames 
metadat <- metadat[order(metadat$Sample),]
dat <- dat[order(dat$Sample),]

# Add the extra variable column
dat$Tube <- metadat$Tube


# Create the Absolute abundance variable 
# Transform Bac_load to Bac_counts
dat$Bac_counts <- 10^dat$Bac_load

# Print the first few rows to check the transformation
head(dat)

# Create absolute abundance variable
dat$Abundance_abs <- dat$Abundance * dat$Bac_counts

# Add a pseudocount of +1 to Abundance_abs to handle zeros before log transforming
# dat$Abundance_log <- log(dat$Abundance_abs + 1) #only if data contains zeros
dat$Abundance_log <- log10(dat$Abundance_abs )


# Turn relevant variables into factors and make sure data is in correct order 
dat <- dat %>% mutate(Treatment = factor(Treatment, levels = c("Saline", "Imolope_low", "Imolope_high"),ordered = TRUE))
dat <- dat %>% mutate_at("Tube", factor)
dat <- dat %>% mutate_at("Tube_ID", factor)
dat <- dat %>% mutate_at("Time_point", factor)

# Create subsets 
dat <- dat[dat$Time_point %in% c('T24', 'T48', 'T72'),]
dat_sal <- dat[dat$Treatment %in% c('Saline'),]
dat_low <- dat[dat$Treatment %in% c('Imolope_low'),] 
dat_high <- dat[dat$Treatment %in% c('Imolope_high'),] 

# Order data sets 
dat_sal <- dat_sal[order(dat_sal$Tube),]
dat_low <- dat_low[order(dat_low$Tube),]
dat_high <- dat_high[order(dat_high$Tube),]

dat_sal <- dat_sal[order(dat_sal$Time_point),]
dat_low <- dat_low[order(dat_low$Time_point),]
dat_high <- dat_high[order(dat_high$Time_point),]
  

# Test normality 
dat %>%
  ggqqplot("Abundance") + facet_grid(as.formula(paste("Time_point","Treatment",sep = " ~ ")))

dat %>%
  group_by(!!sym("Time_point"), !!sym("Treatment")) %>%
  shapiro_test(!!sym("Abundance"))

# Normality assumption is met 

```

#### Visualization + summary statistics 
```{r Bacteroidaceae_vis, eval = TRUE}

# Get summary statistics 
dat %>% 
  group_by(Treatment, Time_point) %>% 
  get_summary_stats(Abundance_log, type = "common")

my_cols_sal <- c("#fbb4ae", "#ff6666", "#CC3333", "#CC0000")
my_cols_low<- c("#99CCFF","#6699CC", "#0066CC","#003399")
my_cols_high <- c("#e9dbf9", "#c994c7","#88419d", "#663366")

plot_sal <- dat_sal %>%
  ggboxplot(x = "Time_point",
           y = "Abundance_log",
           fill = "Time_point") + 
  #geom_line(aes(group = "Tube"), color = "#88419d") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(7, 12), breaks = scales::pretty_breaks(n=4))+
  ylab("Absolute abundance (log)")+
  theme_classic()+
  ggtitle("Saline")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank())+
  scale_fill_manual(values = my_cols_sal)+
stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

plot_low <- dat_low %>%
  ggboxplot(x = "Time_point",
           y = "Abundance_log",
           fill = "Time_point") + 
  #geom_line(aes(group = "Tube"), color = "#CDC673") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(7, 12), breaks = scales::pretty_breaks(n=4))+
  ylab(NULL)+
  theme_classic()+
  ggtitle("Low imolope")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank(), 
        axis.line.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())+
  scale_fill_manual(values = my_cols_low)+
stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

plot_high <- dat_high %>%
  ggboxplot(x = "Time_point",
           y = "Abundance_log",
           fill = "Time_point") + 
  #geom_line(aes(group = "Tube"), color = "#CD853F") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(7, 12), breaks = scales::pretty_breaks(n=4))+
  ylab(NULL)+
  theme_classic()+
  ggtitle("High imolope")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank(), 
        axis.line.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())+
  scale_fill_manual(values = my_cols_high)+
stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))


```

#### Statistical testing 
```{r Bacteroidaceae_stat, eval = TRUE}

# Two-way anova for repeated measures 
two.way <- anova_test(data = dat, 
                      dv = !!sym("Abundance_log"), 
                      wid = !!sym("Tube"), 
                      between = !!sym("Treatment"),
                      within = !!sym("Time_point"))
get_anova_table(two.way) # No significant effects 

```


#### Output plot
```{r Bacteroidaceae_output, eval = TRUE}

# Create output plot
My_Theme = theme(plot.title = element_text(size = 10)) 

plot_sal_final <- plot_sal + My_Theme 

plot_low_final <- plot_low + My_Theme 

plot_high_final <- plot_high + My_Theme 

plot_final <- ggarrange(plot_sal_final, plot_low_final, plot_high_final, ncol=3, nrow=1)

plot_final <- annotate_figure(plot_final, top = text_grob("Bacteroidaceae", color = "black", face = "bold"))

filename <- paste0("plots/In_vitro/Bacteroidaceae.png")
plot_final
suppressMessages(ggsave(filename = filename, width = 2000, height = 1400, units = "px", dpi = "print", device = "png", ))

# Clear the environment and release memory
 rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
 invisible(gc())
 
```

### Porphyromonadaceae
#### Prepare dataset 
```{r Porphyromonadaceae_prep, eval = TRUE}

# Load data 
dat <- read_csv("output/FamilyAbundance_invitro.csv")

# Select data 
dat <- dat[dat$Family %in% c('Porphyromonadaceae'),] 
dat <- subset(dat, Treatment %in% c("Saline", "Imolope_low", "Imolope_high"))

# Add a column with tube identifier that I forgot to add in the metadata in the first place 
# Read metadata 
metadat <- read_csv("input/lop_invitro_meta.csv")
metadat <- subset(metadat, Treatment %in% c("Saline", "Imolope_low", "Imolope_high"))

# remove mock and negative control 
metadat <- metadat[!is.na(metadat[,"Tube"]),]

# Sort both data frames 
metadat <- metadat[order(metadat$Sample),]
dat <- dat[order(dat$Sample),]

# Add the extra variable column
dat$Tube <- metadat$Tube


# Create the Absolute abundance variable 
# Transform Bac_load to Bac_counts
dat$Bac_counts <- 10^dat$Bac_load

# Print the first few rows to check the transformation
head(dat)

# Create absolute abundance variable
dat$Abundance_abs <- dat$Abundance * dat$Bac_counts

# Add a pseudocount of +1 to Abundance_abs to handle zeros before log transforming
# dat$Abundance_log <- log(dat$Abundance_abs + 1) #only if data contains zeros
dat$Abundance_log <- log10(dat$Abundance_abs )



# Turn relevant variables into factors and make sure data is in correct order 
dat <- dat %>% mutate(Treatment = factor(Treatment, levels = c("Saline", "Imolope_low", "Imolope_high"),ordered = TRUE))
dat <- dat %>% mutate_at("Tube", factor)
dat <- dat %>% mutate_at("Tube_ID", factor)
dat <- dat %>% mutate_at("Time_point", factor)

# Create subsets 
dat <- dat[dat$Time_point %in% c('T24', 'T48', 'T72'),]

# Drop unused levels
dat$Time_point <- droplevels(dat$Time_point)

dat_sal <- dat[dat$Treatment %in% c('Saline'),]
dat_low <- dat[dat$Treatment %in% c('Imolope_low'),] 
dat_high <- dat[dat$Treatment %in% c('Imolope_high'),] 

# Order data sets 
dat_sal <- dat_sal[order(dat_sal$Tube),]
dat_low <- dat_low[order(dat_low$Tube),]
dat_high <- dat_high[order(dat_high$Tube),]

dat_sal <- dat_sal[order(dat_sal$Time_point),]
dat_low <- dat_low[order(dat_low$Time_point),]
dat_high <- dat_high[order(dat_high$Time_point),]
  

# Test normality 
dat %>%
  ggqqplot("Abundance") + facet_grid(as.formula(paste("Time_point","Treatment",sep = " ~ ")))

dat %>%
  group_by(!!sym("Time_point"), !!sym("Treatment")) %>%
  shapiro_test(!!sym("Abundance"))

```

#### Visualization + summary statistics 
```{r Porphyromonadaceae_vis, eval = TRUE}

# Get summary statistics 
dat %>% 
  group_by(Treatment, Time_point) %>% 
  get_summary_stats(Abundance_log, type = "common")

my_cols_sal <- c("#fbb4ae", "#ff6666", "#CC3333", "#CC0000")
my_cols_low<- c("#99CCFF","#6699CC", "#0066CC","#003399")
my_cols_high <- c("#e9dbf9", "#c994c7","#88419d", "#663366")

plot_sal <- dat_sal %>%
  ggboxplot(x = "Time_point",
           y = "Abundance_log",
           fill = "Time_point") + 
  #geom_line(aes(group = "Tube"), color = "#88419d") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(9, 11), breaks = scales::pretty_breaks(n=3))+
  ylab("Absolute abundance (log)")+
  theme_classic()+
  ggtitle("Saline")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank())+
  scale_fill_manual(values = my_cols_sal)+
stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

plot_low <- dat_low %>%
  ggboxplot(x = "Time_point",
           y = "Abundance_log",
           fill = "Time_point") + 
  #geom_line(aes(group = "Tube"), color = "#CDC673") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(9, 11), breaks = scales::pretty_breaks(n=3))+
  ylab(NULL)+
  theme_classic()+
  ggtitle("Low imolope")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank(), 
        axis.line.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())+
  scale_fill_manual(values = my_cols_low)+
stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

plot_high <- dat_high %>%
  ggboxplot(x = "Time_point",
           y = "Abundance_log",
           fill = "Time_point") + 
  #geom_line(aes(group = "Tube"), color = "#CD853F") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(9, 11), breaks = scales::pretty_breaks(n=3))+
  ylab(NULL)+
  theme_classic()+
  ggtitle("High imolope")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank(), 
        axis.line.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())+
  scale_fill_manual(values = my_cols_high)+
stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))


```

#### Statistical testing 
```{r Porphyromonadaceae_stat, eval = TRUE}

# Two-way anova for repeated measures 
two.way <- anova_test(data = dat, 
                      dv = !!sym("Abundance_log"), 
                      wid = !!sym("Tube"), 
                      between = !!sym("Treatment"),
                      within = !!sym("Time_point"))
get_anova_table(two.way) # No significance 


```


#### Output plot
```{r Porphyromonadaceae_output, eval = TRUE}

# Create output plot
My_Theme = theme(plot.title = element_text(size = 10)) 

plot_sal_final <- plot_sal + My_Theme 

plot_low_final <- plot_low + My_Theme  

plot_high_final <- plot_high + My_Theme  

plot_final <- ggarrange(plot_sal_final, plot_low_final, plot_high_final, ncol=3, nrow=1)

plot_final <- annotate_figure(plot_final, top = text_grob("Porphyromonadaceae", color = "black", face = "bold"))

filename <- paste0("plots/In_vitro/Porphyromonadaceae.png")
plot_final
suppressMessages(ggsave(filename = filename, width = 2000, height = 1400, units = "px", dpi = "print", device = "png", ))

# Clear the environment and release memory
 rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
 invisible(gc())
 
```
### Erysipelotrichaceae
#### Prepare dataset 
```{r Erysipelotrichaceae_prep, eval = TRUE}
# Load data 
dat <- read_csv("output/FamilyAbundance_invitro.csv")

# Select data 
dat <- dat[dat$Family %in% c('Erysipelotrichaceae'),] 
dat <- subset(dat, Treatment %in% c("Saline", "Imolope_low", "Imolope_high"))

# Add a column with tube identifier that I forgot to add in the metadata in the first place 
# Read metadata 
metadat <- read_csv("input/lop_invitro_meta.csv")
metadat <- subset(metadat, Treatment %in% c("Saline", "Imolope_low", "Imolope_high"))

# remove mock and negative control 
metadat <- metadat[!is.na(metadat[,"Tube"]),]

# Sort both data frames 
metadat <- metadat[order(metadat$Sample),]
dat <- dat[order(dat$Sample),]

# Add the extra variable column
dat$Tube <- metadat$Tube


# Create the Absolute abundance variable 
# Transform Bac_load to Bac_counts
dat$Bac_counts <- 10^dat$Bac_load

# Print the first few rows to check the transformation
head(dat)

# Create absolute abundance variable
dat$Abundance_abs <- dat$Abundance * dat$Bac_counts

# Add a pseudocount of +1 to Abundance_abs to handle zeros before log transforming
# dat$Abundance_log <- log(dat$Abundance_abs + 1) #only if data contains zeros
dat$Abundance_log <- log10(dat$Abundance_abs )



# Turn relevant variables into factors and make sure data is in correct order 
dat <- dat %>% mutate(Treatment = factor(Treatment, levels = c("Saline", "Imolope_low", "Imolope_high"),ordered = TRUE))
dat <- dat %>% mutate_at("Tube", factor)
dat <- dat %>% mutate_at("Tube_ID", factor)
dat <- dat %>% mutate_at("Time_point", factor)

# Create subsets 
dat <- dat[dat$Time_point %in% c('T24', 'T48', 'T72'),]

# Drop unused levels
dat$Time_point <- droplevels(dat$Time_point)

dat_sal <- dat[dat$Treatment %in% c('Saline'),]
dat_low <- dat[dat$Treatment %in% c('Imolope_low'),] 
dat_high <- dat[dat$Treatment %in% c('Imolope_high'),] 

# Order data sets 
dat_sal <- dat_sal[order(dat_sal$Tube),]
dat_low <- dat_low[order(dat_low$Tube),]
dat_high <- dat_high[order(dat_high$Tube),]

dat_sal <- dat_sal[order(dat_sal$Time_point),]
dat_low <- dat_low[order(dat_low$Time_point),]
dat_high <- dat_high[order(dat_high$Time_point),]
  

# Test normality 
dat %>%
  ggqqplot("Abundance") + facet_grid(as.formula(paste("Time_point","Treatment",sep = " ~ ")))

dat %>%
  group_by(!!sym("Time_point"), !!sym("Treatment")) %>%
  shapiro_test(!!sym("Abundance"))

```

#### Visualization + summary statistics 
```{r Erysipelotrichaceae_vis, eval = TRUE}

# Get summary statistics 
dat %>% 
  group_by(Treatment, Time_point) %>% 
  get_summary_stats(Abundance_log, type = "common")

my_cols_sal <- c("#fbb4ae", "#ff6666", "#CC3333", "#CC0000")
my_cols_low<- c("#99CCFF","#6699CC", "#0066CC","#003399")
my_cols_high <- c("#e9dbf9", "#c994c7","#88419d", "#663366")

plot_sal <- dat_sal %>%
  ggboxplot(x = "Time_point",
           y = "Abundance_log",
           fill = "Time_point") + 
  #geom_line(aes(group = "Tube"), color = "#88419d") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(7, 9.5), breaks = scales::pretty_breaks(n=4))+
  ylab("Absolute abundance (log)")+
  theme_classic()+
  ggtitle("Saline")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank())+
  scale_fill_manual(values = my_cols_sal)+
stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

plot_low <- dat_low %>%
  ggboxplot(x = "Time_point",
           y = "Abundance_log",
           fill = "Time_point") + 
  #geom_line(aes(group = "Tube"), color = "#CDC673") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(7, 9.5), breaks = scales::pretty_breaks(n=4))+
  ylab(NULL)+
  theme_classic()+
  ggtitle("Low imolope")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank(), 
        axis.line.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())+
  scale_fill_manual(values = my_cols_low)+
stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

plot_high <- dat_high %>%
  ggboxplot(x = "Time_point",
           y = "Abundance_log",
           fill = "Time_point") + 
  #geom_line(aes(group = "Tube"), color = "#CD853F") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(7, 9.5), breaks = scales::pretty_breaks(n=4))+
  ylab(NULL)+
  theme_classic()+
  ggtitle("High imolope")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank(), 
        axis.line.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())+
  scale_fill_manual(values = my_cols_high)+
stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))


```

#### Statistical testing 
```{r Erysipelotrichaceae_stat, eval = TRUE}

# Two-way anova for repeated measures 
two.way <- anova_test(data = dat, 
                      dv = !!sym("Abundance_log"), 
                      wid = !!sym("Tube"), 
                      between = !!sym("Treatment"),
                      within = !!sym("Time_point"))
get_anova_table(two.way) # No significance 

```


#### Output plot
```{r Erysipelotrichaceae_output, eval = TRUE}

# Create output plot
My_Theme = theme(plot.title = element_text(size = 10)) 

plot_sal_final <- plot_sal + My_Theme 

plot_low_final <- plot_low + My_Theme  

plot_high_final <- plot_high + My_Theme  

plot_final <- ggarrange(plot_sal_final, plot_low_final, plot_high_final, ncol=3, nrow=1)

plot_final <- annotate_figure(plot_final, top = text_grob("Erysipelotrichaceae", color = "black", face = "bold"))

filename <- paste0("plots/In_vitro/Erysipelotrichaceae.png")
plot_final
suppressMessages(ggsave(filename = filename, width = 2000, height = 1400, units = "px", dpi = "print", device = "png", ))

# Clear the environment and release memory
 rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
 invisible(gc())
 
```


### Akkermansiaceae
##### Prepare dataset 
```{r Akkermansiaceae_prep, eval = TRUE}

# Load data 
dat <- read_csv("output/FamilyAbundance_invitro.csv")

# Select data 
dat <- dat[dat$Family %in% c('Akkermansiaceae'),] 
dat <- subset(dat, Treatment %in% c("Saline", "Imolope_low", "Imolope_high"))

# Add a column with tube identifier that I forgot to add in the metadata in the first place 
# Read metadata 
metadat <- read_csv("input/lop_invitro_meta.csv")
metadat <- subset(metadat, Treatment %in% c("Saline", "Imolope_low", "Imolope_high"))

# remove mock and negative control 
metadat <- metadat[!is.na(metadat[,"Tube"]),]

# Sort both data frames 
metadat <- metadat[order(metadat$Sample),]
dat <- dat[order(dat$Sample),]

# Add the extra variable column
dat$Tube <- metadat$Tube


# Create the Absolute abundance variable 
# Transform Bac_load to Bac_counts
dat$Bac_counts <- 10^dat$Bac_load

# Print the first few rows to check the transformation
head(dat)

# Create absolute abundance variable
dat$Abundance_abs <- dat$Abundance * dat$Bac_counts

# Add a pseudocount of +1 to Abundance_abs to handle zeros before log transforming
dat$Abundance_log <- log10(dat$Abundance_abs + 1) #only if data contains zeros


# Turn relevant variables into factors and make sure data is in correct order 
dat <- dat %>% mutate(Treatment = factor(Treatment, levels = c("Saline", "Imolope_low", "Imolope_high"),ordered = TRUE))
dat <- dat %>% mutate_at("Tube", factor)
dat <- dat %>% mutate_at("Tube_ID", factor)
dat <- dat %>% mutate_at("Time_point", factor)

# Create subsets 
dat <- dat[dat$Time_point %in% c('T24', 'T48', 'T72'),]

# Drop unused levels
dat$Time_point <- droplevels(dat$Time_point)

dat_sal <- dat[dat$Treatment %in% c('Saline'),]
dat_low <- dat[dat$Treatment %in% c('Imolope_low'),] 
dat_high <- dat[dat$Treatment %in% c('Imolope_high'),] 

# Order data sets 
dat_sal <- dat_sal[order(dat_sal$Tube),]
dat_low <- dat_low[order(dat_low$Tube),]
dat_high <- dat_high[order(dat_high$Tube),]

dat_sal <- dat_sal[order(dat_sal$Time_point),]
dat_low <- dat_low[order(dat_low$Time_point),]
dat_high <- dat_high[order(dat_high$Time_point),]
  

# Test normality 
dat %>%
  ggqqplot("Abundance_log") + facet_grid(as.formula(paste("Time_point","Treatment",sep = " ~ ")))

dat %>%
  group_by(!!sym("Time_point"), !!sym("Treatment")) %>%
  shapiro_test(!!sym("Abundance_log")) # not working - can't handle zeros? treated as not significant  

```

##### Visualization + summary statistics 
```{r Akkermansiaceae_vis, eval = TRUE}

# Get summary statistics 
dat %>% 
  group_by(Treatment, Time_point) %>% 
  get_summary_stats(Abundance_log, type = "common")

my_cols_sal <- c("#fbb4ae", "#ff6666", "#CC3333", "#CC0000")
my_cols_low<- c("#99CCFF","#6699CC", "#0066CC","#003399")
my_cols_high <- c("#e9dbf9", "#c994c7","#88419d", "#663366")

plot_sal <- dat_sal %>%
  ggboxplot(x = "Time_point",
           y = "Abundance_log",
           fill = "Time_point") + 
  #geom_line(aes(group = "Tube"), color = "#88419d") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(0,7), breaks = scales::pretty_breaks(n=4))+
  ylab("Absolute abundance (log)")+
  theme_classic()+
  ggtitle("Saline")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank())+
  scale_fill_manual(values = my_cols_sal)+
stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

plot_low <- dat_low %>%
  ggboxplot(x = "Time_point",
           y = "Abundance_log",
           fill = "Time_point") + 
  #geom_line(aes(group = "Tube"), color = "#CDC673") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(0,7), breaks = scales::pretty_breaks(n=4))+
  ylab(NULL)+
  theme_classic()+
  ggtitle("Low imolope")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank(), 
        axis.line.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())+
  scale_fill_manual(values = my_cols_low)+
stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

plot_high <- dat_high %>%
  ggboxplot(x = "Time_point",
           y = "Abundance_log",
           fill = "Time_point") + 
  #geom_line(aes(group = "Tube"), color = "#CD853F") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(0,7), breaks = scales::pretty_breaks(n=4))+
  ylab(NULL)+
  theme_classic()+
  ggtitle("High imolope")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank(), 
        axis.line.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())+
  scale_fill_manual(values = my_cols_high)+
stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))


```

##### Statistical testing 
```{r Akkermansiaceae_stat, eval = TRUE}

# Friedman test 
f <- friedman.test(Abundance_log ~ Time_point|Tube, data = dat) # Significant 

# Post hoc tests 
stat.test.sal <- dat_sal %>% 
  wilcox_test(Abundance_log~Time_point, paired = TRUE, p.adjust.method = "fdr") %>%
  add_significance("p.adj") %>%
  add_xy_position(x = "Time_point", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE)

stat.test.low <- dat_low %>% 
  wilcox_test(Abundance_log~Time_point, paired = TRUE, p.adjust.method = "fdr") %>%
  add_significance("p.adj") %>%
  add_xy_position(x = "Time_point", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE)%>%
 filter(!is.na(p.adj))

stat.test.high <- dat_high %>% 
  wilcox_test(Abundance_log~Time_point, paired = TRUE, p.adjust.method = "fdr") %>%
  add_significance("p.adj") %>%
  add_xy_position(x = "Time_point", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE)
```


##### Output plot
```{r Akkermansiaceae_output, eval = TRUE}

# Create output plot
My_Theme = theme(plot.title = element_text(size = 10)) 

plot_sal_final <- plot_sal + My_Theme + 
  stat_pvalue_manual(stat.test.sal, label = "p.adj.signif", tip.length = 0.01, size = 3.5, hide.ns = TRUE)

plot_low_final <- plot_low + My_Theme + 
  stat_pvalue_manual(stat.test.low, label = "p.adj.signif", tip.length = 0.01, size = 3.5, hide.ns = TRUE) 

plot_high_final <- plot_high + My_Theme +  
  stat_pvalue_manual(stat.test.high, label = "p.adj.signif", tip.length = 0.01, size = 3.5, hide.ns = TRUE) 

plot_final <- ggarrange(plot_sal_final, plot_low_final, plot_high_final, ncol=3, nrow=1)

plot_final <- annotate_figure(plot_final, top = text_grob("Akkermansiaceae", color = "black", face = "bold"))

filename <- paste0("plots/In_vitro/Akkermansiaceae_trajectories.png")
plot_final
suppressMessages(ggsave(filename = filename, width = 2000, height = 1400, units = "px", dpi = "print", device = "png", ))

# Clear the environment and release memory
 rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
 invisible(gc())
 
```

# ***********************
##NOT INCLUDED IN MANUSCRIPT

### BAC LOAD 
#### Timepoint 24h
##### Prepare dataset 
```{r load_24h_prep, eval = TRUE}

# Load data 
dat <- read_csv("input/lop_invitro_meta.csv")

# Select data 
dat <- dat[dat$Time_point%in% c('T24'),]
dat <- subset(dat, Treatment %in% c("Saline", "Imolope_low", "Imolope_high"))


# Turn relevant variables into factors and make sure data is in correct order 
dat <- dat %>% mutate(Treatment = factor(Treatment, levels = c("Saline", "Imolope_low", "Imolope_high"),ordered = TRUE))
dat <- dat %>% mutate_at("Tube_ID", factor)
dat <- dat %>% mutate_at("Time_point", factor)
  
# Test normality 
shapiro_test(dat,Bac_load)  
ggqqplot(dat,x = "Bac_load") # Normality: yes


```

##### Visualization + summary statistics 
```{r load_24h_vis, eval = TRUE}

# Get summary statistics 
dat %>% 
  group_by(Treatment) %>% 
  get_summary_stats(Bac_load, type = "common")


# Visualise data
plot <- dat %>%
  ggboxplot(x = "Treatment",
           y = "Bac_load",
           fill = "Treatment") + 
  geom_point(size = 1, color = "black") +
  #scale_y_continuous(limits=c(9,12))+
  ylab("16S copies/ul (log)")+
  ggtitle("Bacterial load (24h)")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        legend.position = "right")+ 
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))+
  scale_fill_manual(values=c("#CC3333","#0066CC", "#88419d"), labels = c("Saline vehicle", "Saline + low imolope", "Saline + high imolope"))




```

##### Statistical testing 
```{r load_24h_stat, eval = TRUE}

# ANOVA 
anova <- anova_test(Bac_load ~ Treatment, data = dat) # not significant


print(plot)
ggsave(filename = "plots/In_vitro/Bac_load_t24.png", plot = plot,
       dpi = "print", height = 1400, width = 1800, units = "px", device = "png")


# Clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```

#### Timepoint 48h 
##### Prepare dataset 
```{r load_48h_prep, eval = TRUE}

# Load data 
dat <- read_csv("input/lop_invitro_meta.csv")

# Select data 
dat <- dat[dat$Time_point%in% c('T48'),] 
dat <- subset(dat, Treatment %in% c("Saline", "Imolope_low", "Imolope_high"))


# Turn relevant variables into factors and make sure data is in correct order 
dat <- dat %>% mutate(Treatment = factor(Treatment, levels = c("Saline", "Imolope_low", "Imolope_high"),ordered = TRUE))
dat <- dat %>% mutate_at("Tube_ID", factor)
dat <- dat %>% mutate_at("Time_point", factor)
  
# Test normality 
shapiro_test(dat,Bac_load)  
ggqqplot(dat,x = "Bac_load") # Normality: yes


```

##### Visualization + summary statistics 
```{r load_48h_vis, eval = TRUE}

# Get summary statistics 
dat %>% 
  group_by(Treatment) %>% 
  get_summary_stats(Bac_load, type = "common")


# Visualise data
plot <- dat %>%
  ggboxplot(x = "Treatment",
           y = "Bac_load",
           fill = "Treatment") + 
  geom_point(size = 1, color = "black") +
  #scale_y_continuous(limits=c(9,12))+
  ylab("16S copies/ul (log)")+
  ggtitle("Bacterial load (48h)")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        legend.position = "right")+ 
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))+
  scale_fill_manual(values=c("#CC3333","#0066CC", "#88419d"), labels = c("Saline vehicle", "Saline + low imolope", "Saline + high imolope"))




```

##### Statistical testing 
```{r load_48h_stat, eval = TRUE}

# ANOVA 
anova <- anova_test(Bac_load ~ Treatment, data = dat) # significant

# Post hoc test 
stat.test <- dat %>% 
  t_test(Bac_load~Treatment, p.adjust.method = "fdr") %>% 
  add_significance("p.adj") %>%
  add_xy_position(x = "Treatment", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE) # nothing significant after correction for multiple comparisons 


print(plot)
ggsave(filename = "plots/In_vitro/Bac_load_t48.png", plot = plot,
       dpi = "print", height = 1400, width = 1800, units = "px", device = "png")


# Clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```

#### Timepoint 72h 
##### Prepare dataset 
```{r load_72h_prep, eval = TRUE}

# Load data 
dat <- read_csv("input/lop_invitro_meta.csv")

# Select data 
dat <- dat[dat$Time_point%in% c('T72'),] 
dat <- subset(dat, Treatment %in% c("Saline", "Imolope_low", "Imolope_high"))


# Turn relevant variables into factors and make sure data is in correct order 
dat <- dat %>% mutate(Treatment = factor(Treatment, levels = c("Saline", "Imolope_low", "Imolope_high"),ordered = TRUE))
dat <- dat %>% mutate_at("Tube_ID", factor)
dat <- dat %>% mutate_at("Time_point", factor)
  
# Test normality 
shapiro_test(dat,Bac_load)  
ggqqplot(dat,x = "Bac_load") # Normality: yes


```

##### Visualization + summary statistics 
```{r load_72h_vis, eval = TRUE}

# Get summary statistics 
dat %>% 
  group_by(Treatment) %>% 
  get_summary_stats(Bac_load, type = "common")


# Visualise data
plot <- dat %>%
  ggboxplot(x = "Treatment",
           y = "Bac_load",
           fill = "Treatment") + 
  geom_point(size = 1, color = "black") +
  #scale_y_continuous(limits=c(9,12))+
  ylab("16S copies/ul (log)")+
  ggtitle("Bacterial load (72h)")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        legend.position = "right")+ 
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))+
  scale_fill_manual(values=c("#CC3333","#0066CC", "#88419d"), labels = c("Saline vehicle", "Saline + low imolope", "Saline + high imolope"))




```

##### Statistical testing 
```{r load_72h_stat, eval = TRUE}

# ANOVA 
anova <- anova_test(Bac_load ~ Treatment, data = dat) # not significant


print(plot)
ggsave(filename = "plots/In_vitro/Bac_load_t72.png", plot = plot,
       dpi = "print", height = 1400, width = 1800, units = "px", device = "png")


# Clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```


