---
title: "PRIMA06 Metadata analysis"
author: "annapiih"
date: '`r Sys.Date()`'
output:
  rmdformats::readthedown:
    toc_depth: 4
    collapsed: false
    code_folding: hide
    number_sections: true
knit: (function(inputFile, encoding) { 
    rmarkdown::render(
        inputFile, encoding = encoding,
        output_dir = file.path(dirname(inputFile),"output"),
        output_file = paste0("PRIMA06_", Sys.Date(), "_Animals.html")) 
    })
---

# Intro {.tabset .tabset-fade .tabset-pills}

This markdown contains commands for analyzing metadata in animal studies (transit time, body weight, food intake). 

By default, R uses the folder where the markdown file is saved as the working directory  


```{r setup, eval=TRUE, echo=TRUE, message=FALSE,warning = FALSE}
knitr::opts_chunk$set(echo = TRUE ,warning = FALSE, message = FALSE)

# Load libraries
library(tidyverse)
library(phyloseq)
library(ggpubr)
library(rstatix)
library(kableExtra)
library(picante)
library(readr)
library(forcats)
library(ggsci)
library (RColorBrewer)
library(viridis)
library(ggpmisc)
library(multipanelfigure)
library(ggplot2)
library(grid)

# Create used folders if missing
if (!file.exists("plots")) dir.create(file.path(getwd(), "plots"))
if (!file.exists("scripts")) dir.create(file.path(getwd(), "scripts"))
if (!file.exists("output")) dir.create(file.path(getwd(), "output"))

```


# TT between-group comparison (change in minutes)

## Prepare dataset 
```{r TT_b_prep, eval = TRUE}
# Load data 
metadat <- read_csv("input/input_PRIMA06_animals.csv")

# Create the dataset 
dat <- metadat[metadat$Day %in% c('D05','D08','D10'),] 

dat$Day

# Turn relevant variables into factors and make sure data is in correct order 
dat <- dat %>% mutate_at("Group", factor, levels = c("Control", "Low", "Medium", "High")) %>% mutate_at("Day", factor, levels = c("D05", "D08", "D10")) %>% mutate_at("Animal", factor) 

# Make sure that data is in correct order (for paired analysis)
dat <- dat[order(dat$Day),]


```

## Visualization + summary statistics 
```{r TT_b_vis, eval = TRUE}

# Plot 
plot <-
  ggboxplot(dat, x = "Group", y = "TT_Increase",
              fill = "Group", palette = "Pastel1", outlier.shape = NA) + 
  ggtitle("Change in transit time from baseline") + 
  facet_wrap("Day", nrow = 1, scales = "fixed") +
  ylab("Minutes change from baseline (day 2)")+ 
  #theme(plot.title = element_text(hjust = 0.5)) +
  theme_classic()+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  geom_point(aes(shape = Group), 
             size = 1.5, 
             position = position_jitterdodge(jitter.width = 0.01, dodge.width = 0.8)) + 
  scale_shape_manual(values=c(20,20,20,20))+
  scale_y_continuous(limits=c(-150, 700), breaks = scales::pretty_breaks(n=7))+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))


# Get summary statistics 
dat %>% 
  group_by(Group, Day) %>% 
  get_summary_stats(TT_Increase, type = "common")

```

## Check assumptions 
```{r TT_b_as, eval = TRUE}

#Shapiro test 
shapiro_test(dat, TT_Increase) 

ggqqplot(dat,x = "TT_Increase")  


# Levene's test (test for homogeneity)
dat %>% 
  levene_test(TT_Increase~Group*Day)

```

## Statistical testing 
```{r TT_b_stat, eval = TRUE}

# Two-way ANOVA 
two.way <- aov(TT_Increase ~ Day*Group + Error(Animal), data = dat)
summary(two.way)


# Post hoc test 2 
stat.test <- dat %>% 
  group_by(Day)%>% 
  t_test(TT_Increase~Group, p.adjust.method = "fdr") %>% 
  add_significance("p.adj") %>%
  add_xy_position(x = "Group", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE)


plot_final <- plot + 
  stat_pvalue_manual(stat.test, 
                     label = "p.adj.signif", 
                     tip.length = 0.01, 
                     hide.ns = TRUE, 
                     size = 3.5, 
                     y.position = c(560,600,640,600,520,560,600,640)) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

filename <- paste0("plots/Animals/TT_between.png")
plot_final
suppressMessages(ggsave(filename = filename, width = 2400, height = 1400, units = "px", dpi = "print", device = "png", ))


# Clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```

# TT within-group comparison (change in minutes)
## Prepare dataset 
```{r TT_w_prep, eval = TRUE}
# Load data 
metadat <- read_csv("input/input_PRIMA06_animals.csv")

# Create the dataset 
dat_tt2 <- metadat[metadat$Day %in% c('D05','D08','D10'),] 

dat_tt2$Day

# Turn relevant variables into factors and make sure data is in correct order 
dat_tt2 <- dat_tt2 %>% mutate_at("Group", factor, levels = c("Control", "Low", "Medium", "High")) %>% mutate_at("Day", factor, levels = c("D05", "D08", "D10")) %>% mutate_at("Animal", factor) 

# Make sure that data is in correct order (for paired analysis)
dat_tt2 <- dat_tt2[order(dat_tt2$Day),]

```


## Visualization + summary statistics 
```{r TT_w_vis, eval = TRUE}

# Plot 
plot <-
  ggboxplot(dat_tt2, x = "Day", y = "TT_Increase",
              fill = "Group", palette = "Pastel1", outlier.shape = NA) + 
  ggtitle("Change in transit time from baseline") + 
  facet_wrap("Group", nrow = 1, scales = "fixed") +
  ylab("Minutes change from baseline (day 2)")+ 
  #theme(plot.title = element_text(hjust = 0.5)) +
  theme_classic()+
  geom_point(aes(shape = Group), 
             size = 1.5, 
             position = position_jitterdodge(jitter.width = 0.01, dodge.width = 0.8)) + 
  scale_shape_manual(values=c(20,20,20,20))+
  scale_y_continuous(limits=c(-150, 700), breaks = scales::pretty_breaks(n=7))+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

# Get summary statistics 
dat_tt2 %>% 
  group_by(Group, Day) %>% 
  get_summary_stats(TT_Increase, type = "common")

```

## Check assumptions 
```{r TT_w_as, eval = TRUE}

#Shapiro test 
shapiro_test(dat_tt2, TT_Increase) 

ggqqplot(dat_tt2,x = "TT_Increase") # Normal distributed   


# Levene's test (test for homogeneity)
dat_tt2 %>% 
  levene_test(TT_Increase~Group*Day) # Assumption met 

```

## Statistical testing 
```{r TT_w_stat, eval = TRUE}

# Two-way anova for repeated measures 
two.way <- aov(TT_Increase ~ Day*Group + Error(Animal), data = dat_tt2)
summary(two.way)

# Post hoc test  
stat.test2 <- dat_tt2 %>% 
  group_by(Group)%>% 
  t_test(TT_Increase~Day, paired = TRUE, p.adjust.method = "fdr") %>% 
  add_significance("p.adj") %>%
  add_xy_position(x = "Day", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE)

plot_tt2_final <- plot+
  stat_pvalue_manual(stat.test2, 
                     label = "p.adj.signif", 
                     tip.length = 0.01, 
                     hide.ns = TRUE, 
                     size = 3.5) 

filename <- paste0("plots/Animals/TT_within.png")
plot_tt2_final
suppressMessages(ggsave(filename = filename, width = 2400, height = 1400, units = "px", dpi = "print", device = "png", ))


# Clear the environment and release memory
 rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
 invisible(gc())

```


# BW 

## Prepare data 
```{r Prepare_BW, eval = TRUE}

# Load data 
metadat <- read_csv("input/input_PRIMA06_animals.csv")

# Remove samples with incomplete metadata (i.e., remove days where transit time was not measured)
dat_BW <- metadat[metadat$Day %in% c('D02','D05','D08','D10'),] 
dat_BW$Day 

# Turn relevant variables into factors and make sure data is in correct order 
dat_BW <- dat_BW %>% mutate_at("Group", factor, levels = c("Control", "Low", "Medium", "High")) %>% mutate_at("Day", factor, levels = c("D02", "D05", "D08", "D10")) %>% mutate_at("Animal", factor) 

# Make sure that data is in correct order (for paired analysis)
dat_BW <- dat_BW[order(dat_BW$Day),]

dat_BW_con <- dat_BW[dat_BW$Group %in% c('Control'),] 
dat_BW_low <- dat_BW[dat_BW$Group %in% c('Low'),] 
dat_BW_med <- dat_BW[dat_BW$Group %in% c('Medium'),] 
dat_BW_high <- dat_BW[dat_BW$Group %in% c('High'),] 

```

## Visualization + summary statistics 
```{r Visualize_BW, eval = TRUE}

# Get summary statistics 
dat_BW %>% 
  group_by(Group, Day) %>% 
  get_summary_stats(BW, type = "common")

# Set colors 
my_cols_con <- c("#fbb4ae", "#ff6666", "#CC3333", "#CC0000")
my_cols_low <- c("#99CCFF","#6699CC", "#0066CC","#003399")
my_cols_med <- c("darkseagreen2", "#66CC66", "#339933", "#006600")
my_cols_high <- c("#e9dbf9", "#c994c7","#88419d", "#663366")

# Visualise data 
plot_con <- dat_BW_con %>%
  ggboxplot(x = "Day",
           y = "BW",
           fill = "Day") + 
  geom_line(aes(group = Animal), color = "#ff6666") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(18, 23), breaks = scales::pretty_breaks(n=4))+
  ylab("Body weight (g)")+
  theme_classic()+
  ggtitle("Control")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank())+
  scale_fill_manual(values = my_cols_con)+
stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

plot_low <- dat_BW_low %>%
  ggboxplot(x = "Day",
           y = "BW",
           fill = "Day") + 
  geom_line(aes(group = Animal), color = "#3366CC") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(18, 23), breaks = scales::pretty_breaks(n=4))+
  ylab(NULL)+
  theme_classic()+
  ggtitle("Low dose")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank(), 
        axis.line.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())+
  scale_fill_manual(values = my_cols_low)+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

plot_med <- dat_BW_med %>%
  ggboxplot(x = "Day",
           y = "BW",
           fill = "Day") + 
  geom_line(aes(group = Animal), color = "#66CC66") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(18, 23), breaks = scales::pretty_breaks(n=4))+
  ylab(NULL)+
  theme_classic()+
  ggtitle("Medium dose")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank(), 
        axis.line.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())+
  scale_fill_manual(values = my_cols_med)+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

plot_high <- dat_BW_high %>%
  ggboxplot(x = "Day",
           y = "BW",
           fill = "Day") + 
  geom_line(aes(group = Animal), color = "#c994c7") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(18, 23), breaks = scales::pretty_breaks(n=4))+
  ylab(NULL)+
  theme_classic()+
  ggtitle("High dose")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank(), 
        axis.line.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())+
  scale_fill_manual(values = my_cols_high)+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))


```

## Check assumptions 
```{r assumptions_BW, eval = TRUE}

# Shapiro tests 
dat_BW %>%
  group_by(Group)%>%
  shapiro_test(BW)

# QQ plots
dat_BW %>%
  ggqqplot("BW") + facet_grid("Group")


# Levene's test (test for homogeneity)
dat_BW %>% 
  levene_test(BW~Group*Day)

```

## Statistical testing 
```{r stat_BW, eval = TRUE}

# Friedman for paired data 
friedman.test.within <- friedman.test(BW ~ Day|Animal, data = dat_BW)
friedman.test.within


# Within group comparison: Paired Wilcoxon  
stat.test.con <- dat_BW_con %>%
  wilcox_test(as.formula(paste("BW ~", "Day", sep = " ")), paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj") %>% 
  add_xy_position(x = "Day", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE)

stat.test.low <- dat_BW_low %>%
  wilcox_test(as.formula(paste("BW ~", "Day", sep = " ")), paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj") %>% 
  add_xy_position(x = "Day", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE)

stat.test.med <- dat_BW_med %>%
  wilcox_test(as.formula(paste("BW ~", "Day", sep = " ")), paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj") %>% 
  add_xy_position(x = "Day", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE)

stat.test.high <- dat_BW_high %>%
  wilcox_test(as.formula(paste("BW ~", "Day", sep = " ")), paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj") %>% 
  add_xy_position(x = "Day", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE)


# Create output plot
My_Theme = theme(plot.title = element_text(size = 10)) # For setting the text sizes in the plot 

plot_con_final <- plot_con + My_Theme + 
  stat_pvalue_manual(stat.test.con, label = "p.adj.signif", tip.length = 0.01, size = 3.5, hide.ns = TRUE)

plot_low_final <- plot_low + My_Theme + 
  stat_pvalue_manual(stat.test.low, label = "p.adj.signif", tip.length = 0.01, size = 3.5, hide.ns = TRUE) 

plot_med_final <- plot_med + My_Theme + 
  stat_pvalue_manual(stat.test.med, label = "p.adj.signif", tip.length = 0.01, size = 3.5, hide.ns = TRUE) 

plot_high_final <- plot_high + My_Theme +  
  stat_pvalue_manual(stat.test.high, label = "p.adj.signif", tip.length = 0.01, size = 3.5, hide.ns = TRUE) 

plot_final <- ggarrange(plot_con_final, plot_low_final, plot_med_final, plot_high_final, ncol=4, nrow=1)

plot_final <- annotate_figure(plot_final, top = text_grob("Body weight during the experiment", color = "black", face = "bold"))

filename <- paste0("plots/Animals/BW_Within.png")
plot_final
suppressMessages(ggsave(filename = filename, width = 2000, height = 1400, units = "px", dpi = "print", device = "png", ))


# Clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```


# Food intake 

## Prepare data 
```{r prepare_food, eval = TRUE}

# Load data 
metadat <- read_csv("input/input_PRIMA06_animals.csv")

# Find out where data on food intake is missing 
metadat[!complete.cases(metadat[,"FoodIntake"]), ] 
# All animals have missing values for day1 and two animals have missing values for day2 - but it is difficult to leave out day 2 when day1 and day2 are the only days before treatment. Therefore, I will instead remove the animals with missing values from the analysis 

# Create the new dataset 
dat_food <- metadat[metadat$Day %in% c('D02', 'D05','D08','D10'),] 
dat_food <- dat_food[!dat_food$Animal %in% c('M11', 'M23'),]


# Turn relevant variables into factors and make sure data is in correct order 
dat_food <- dat_food %>% mutate_at("Group", factor, levels = c("Control", "Low", "Medium", "High")) %>% mutate_at("Day", factor, levels = c("D02", "D05", "D08", "D10")) %>% mutate_at("Animal", factor) 

# Make sure that data is in correct order (for paired analysis)
dat_food <- dat_food[order(dat_food$Day),]

dat_food_con <- dat_food[dat_food$Group %in% c('Control'),] 
dat_food_low <- dat_food[dat_food$Group %in% c('Low'),]
dat_food_med <- dat_food[dat_food$Group %in% c('Medium'),]
dat_food_high <- dat_food[dat_food$Group %in% c('High'),]

```

## Visualization + summary statistics 
```{r visualize_food, eval = TRUE}


# Get summary statistics 
dat_food %>% 
  group_by(Group, Day) %>% 
  get_summary_stats(FoodIntake, type = "common")

# Set colors 
my_cols_con <- c("#fbb4ae", "#ff6666", "#CC3333", "#CC0000")
my_cols_low <- c("#99CCFF","#6699CC", "#0066CC","#003399")
my_cols_med <- c("darkseagreen2", "#66CC66", "#339933", "#006600")
my_cols_high <- c("#e9dbf9", "#c994c7","#88419d", "#663366")

# Visualise data 
plot_con <- dat_food_con %>%
  ggboxplot(x = "Day",
           y = "FoodIntake",
           fill = "Day") + 
  geom_line(aes(group = Animal), color = "#ff6666") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(0, 6), breaks = scales::pretty_breaks(n=4))+
  ylab("Food intake (g)")+
  theme_classic()+
  ggtitle("Control")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank())+
  scale_fill_manual(values = my_cols_con)+
stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

plot_low <- dat_food_low %>%
  ggboxplot(x = "Day",
           y = "FoodIntake",
           fill = "Day") + 
  geom_line(aes(group = Animal), color = "#3366CC") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(0, 6), breaks = scales::pretty_breaks(n=4))+
  ylab(NULL)+
  theme_classic()+
  ggtitle("Low dose")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank(), 
        axis.line.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())+
  scale_fill_manual(values = my_cols_low)+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

plot_med <- dat_food_med %>%
  ggboxplot(x = "Day",
           y = "FoodIntake",
           fill = "Day") + 
  geom_line(aes(group = Animal), color = "#66CC66") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(0, 6), breaks = scales::pretty_breaks(n=4))+
  ylab(NULL)+
  theme_classic()+
  ggtitle("Medium dose")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank(), 
        axis.line.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())+
  scale_fill_manual(values = my_cols_med)+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

plot_high <- dat_food_high %>%
  ggboxplot(x = "Day",
           y = "FoodIntake",
           fill = "Day") + 
  geom_line(aes(group = Animal), color = "#c994c7") + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(0, 6), breaks = scales::pretty_breaks(n=4))+
  ylab(NULL)+
  theme_classic()+
  ggtitle("High dose")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position="none")+ 
  theme(axis.title.x=element_blank(), 
        axis.line.y = element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())+
  scale_fill_manual(values = my_cols_high)+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))



```

## Check assumptions 
```{r assumptions_food, eval = TRUE}

# Shapiro tests 
dat_food %>%
  group_by(Group)%>%
  shapiro_test(FoodIntake) # not normally distributed 

# QQ plots
dat_food %>%
  ggqqplot("FoodIntake") + facet_grid("Group")


# Levene's test (test for homogeneity)
dat_food %>% 
  levene_test(FoodIntake~Group*Day)

```


## Statistical testing 
```{r stat_food, eval = TRUE}

# Friedman for paired data 
friedman.test.within <- friedman.test(FoodIntake ~ Day|Animal, data = dat_food)
friedman.test.within

# Within group comparison: Paired Wilcoxon  
stat.test.con <- dat_food_con %>%
  wilcox_test(as.formula(paste("FoodIntake ~", "Day", sep = " ")), paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj") %>% 
  add_xy_position(x = "Day", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE)

stat.test.low <- dat_food_low %>%
  wilcox_test(as.formula(paste("FoodIntake ~", "Day", sep = " ")), paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj") %>% 
  add_xy_position(x = "Day", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE)

stat.test.med <- dat_food_med %>%
  wilcox_test(as.formula(paste("FoodIntake ~", "Day", sep = " ")), paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj") %>% 
  add_xy_position(x = "Day", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE)

stat.test.high <- dat_food_high %>%
  wilcox_test(as.formula(paste("FoodIntake ~", "Day", sep = " ")), paired = TRUE) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj") %>% 
  add_xy_position(x = "Day", dodge = 0.8) %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE)


# Create output plot
My_Theme = theme(plot.title = element_text(size = 10)) # For setting the text sizes in the plot 

plot_con_final <- plot_con + My_Theme + 
  stat_pvalue_manual(stat.test.con, label = "p.adj.signif", tip.length = 0.01, size = 3.5, hide.ns = TRUE)

plot_low_final <- plot_low + My_Theme + 
  stat_pvalue_manual(stat.test.low, label = "p.adj.signif", tip.length = 0.01, size = 3.5, hide.ns = TRUE) 

plot_med_final <- plot_med + My_Theme + 
  stat_pvalue_manual(stat.test.med, label = "p.adj.signif", tip.length = 0.01, size = 3.5, hide.ns = TRUE) 

plot_high_final <- plot_high + My_Theme +  
  stat_pvalue_manual(stat.test.high, label = "p.adj.signif", tip.length = 0.01, size = 3.5, hide.ns = TRUE) 

plot_final <- ggarrange(plot_con_final, plot_low_final, plot_med_final, plot_high_final, ncol=4, nrow=1)

plot_final <- annotate_figure(plot_final, top = text_grob("Food intake during the experiment", color = "black", face = "bold"))

filename <- paste0("plots/Animals/Food.png")
plot_final
suppressMessages(ggsave(filename = filename, width = 2000, height = 1400, units = "px", dpi = "print", device = "png", ))


# Clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```
# INFO {.tabset .tabset-fade .tabset-pills}

## FINAL COMMENT

This completes the statistical analysis of animal data in PRIMA06 


## SESSION INFO

The analysis was run in the following environment:

```{r packages, eval=TRUE}
sessionInfo()
```


