---
title: "PRIMA06 Metabolite analysis"
author: "annapiih"
date: '`r Sys.Date()`'
output:
  rmdformats::readthedown:
    toc_depth: 4
    collapsed: false
    code_folding: hide
    number_sections: true
knit: (function(inputFile, encoding) { 
    rmarkdown::render(
        inputFile, encoding = encoding,
        output_dir = file.path(dirname(inputFile),"output"),
        output_file = paste0("PRIMA06_", Sys.Date(), "_Metabolites.html")) 
    })
---

# Intro {.tabset .tabset-fade .tabset-pills}

This markdown contains commands for analysing metabolite data from the animal study PRIMA06. 

By default, R uses the folder where the markdown file is saved as the working directory  


```{r setup, eval=TRUE, echo=TRUE, message=FALSE,warning = FALSE}
knitr::opts_chunk$set(echo = TRUE ,warning = FALSE, message = FALSE)

# Load libraries
library(tidyverse)
library(phyloseq)
library(ggpubr)
library(rstatix)
library(kableExtra)
library(picante)
library(readr)
library(forcats)
library(ggsci)
library (RColorBrewer)
library(viridis)
library(ggpmisc)
library(multipanelfigure)
library(ggplot2)
library(grid)

# Create used folders if missing
if (!file.exists("plots")) dir.create(file.path(getwd(), "plots"))
if (!file.exists("scripts")) dir.create(file.path(getwd(), "scripts"))
if (!file.exists("output")) dir.create(file.path(getwd(), "output"))

```


# Acetic acid (cecum){.tabset .tabset-fade .tabset-pills}

## Prepare dataset 
```{r prepare_acetic, eval = TRUE}

# Load data 
metadat <- read_csv("input/input_PRIMA06_animals.csv")

# Remove samples with incomplete metadata (i.e., remove all days except day 11)
dat_acetic <- metadat[!is.na(metadat[,"Acetic_acid_cecum"]),]

# Turn relevant variables into factors and make sure data is in correct order 
dat_acetic <- dat_acetic %>% mutate_at("Group", factor, levels = c("Control", "Low", "Medium", "High")) %>% mutate_at("Day", factor, levels = c("D11")) %>% mutate_at("Animal", factor) 

# Identify outliers 
out <- boxplot.stats(dat_acetic$Acetic_acid_cecum)$out
out_ind <- which(dat_acetic$Acetic_acid_cecum %in% c(out))
out_ind

dat_acetic[out_ind, ]

# Test normality 
shapiro_test(dat_acetic, Acetic_acid_cecum) # Not normally distributed
ggqqplot(dat_acetic,x = "Acetic_acid_cecum") 

```

## Visualization + summary statistics 
```{r visualize_acetic, eval = TRUE}

# Get summary statistics 
dat_acetic %>% 
  group_by(Group) %>% 
  get_summary_stats(Acetic_acid_cecum, type = "mean_sd")

# Visualize data 
plot_acetic <-ggboxplot(dat_acetic, 
                        x = "Group", 
                        y = "Acetic_acid_cecum",
                        fill = "Group", 
                        palette = "Pastel1", 
                        outlier.shape = NA) + 
  geom_point(size = 0.5, color = "black") +
  ylab("ug/g cecum")+
  xlab(NULL)+
  ggtitle("Acetic acid in cecum")+
  scale_y_continuous(limits=c(2000,9000), breaks = scales::pretty_breaks(n=8))+
  theme_classic()+
  theme(legend.position="none")+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))


```

## Statistical testing 

```{r acetic_stat, eval = TRUE}

# Kruskal wallis for unpaired data 
kruskal.test <- kruskal.test(Acetic_acid_cecum ~ Group, data = dat_acetic) # not significant 


filename <- paste0("plots/Metabolites/acetic_acid_cecum.png")
plot_acetic 
suppressMessages(ggsave(filename = filename, width = 1800, height = 1400, units = "px", dpi = "print", device = "png", ))

# Clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```

# Propanoic acid (cecum) {.tabset .tabset-fade .tabset-pills}

## Prepare dataset 
```{r prepare_propanoic, eval = TRUE}

# Load data 
metadat <- read_csv("input/input_PRIMA06_animals.csv")

# Remove samples with incomplete metadata (i.e., remove all days except day 11)
dat_prop <- metadat[!is.na(metadat[,"Propanoic_acid_cecum"]),]

# Turn relevant variables into factors and make sure data is in correct order
dat_prop <- dat_prop %>% mutate_at("Group", factor, levels = c("Control", "Low", "Medium", "High")) %>% mutate_at("Day", factor, levels = c("D11")) %>% mutate_at("Animal", factor) 

# Identify outliers 
out <- boxplot.stats(dat_prop$Propanoic_acid_cecum)$out
out_ind <- which(dat_prop$Propanoic_acid_cecum %in% c(out))
out_ind

dat_prop[out_ind, ]

# Remove animal 21 and 22 before statistical analysis (high outliers in the high-dose group) 
dat_prop_stat <- dat_prop[-c(21,22),]

# Test normality 
shapiro_test(dat_prop_stat, Propanoic_acid_cecum) # Normally distributed

ggqqplot(dat_prop_stat,x = "Propanoic_acid_cecum") 

# Levene's test (test for homogeneity)
dat_prop_stat %>% 
  levene_test(Propanoic_acid_cecum~Group) # Assumption is met 

```

## Visualization + summary statistics 
```{r visualize_propanoic, eval = TRUE}

# Get summary statistics 
dat_prop %>% 
  group_by(Group) %>% 
  get_summary_stats(Propanoic_acid_cecum, type = "common") 

# Visualize data (with outliers)
plot_prop <-ggboxplot(dat_prop, 
                      x = "Group", 
                      y = "Propanoic_acid_cecum",
                      fill = "Group", 
                      palette = "Pastel1", 
                      outlier.shape = NA) + 
  geom_point(size = 0.5, color = "black") +
  ylab("ug/g cecum")+
  xlab(NULL)+
  ggtitle("Propanoic acid in cecum")+
  #scale_y_continuous(limits=c(450,1450), breaks = scales::pretty_breaks(n=6))+
  theme_classic()+
  theme(legend.position="none")+ 
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

# Visualize data (without outliers)
plot_prop_2 <-ggboxplot(dat_prop_stat, 
                      x = "Group", 
                      y = "Propanoic_acid_cecum",
                      fill = "Group", 
                      palette = "Pastel1", 
                      outlier.shape = NA) + 
  geom_point(size = 0.5, color = "black") +
  ylab("ug/g cecum")+
  xlab(NULL)+
  ggtitle("Propanoic acid in cecum")+
  #scale_y_continuous(limits=c(450,1450), breaks = scales::pretty_breaks(n=6))+
  theme_classic()+
  theme(legend.position="none")+ 
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

```

## Statistical testing 

```{r propanoic_stat, eval = TRUE}

# ANOVA  
anova <- aov(Propanoic_acid_cecum ~ Group, data = dat_prop_stat)
summary(anova)

# Between group comparison: Unpaired Wilcoxon  
stat.test <- dat_prop_stat %>%
  t_test(Propanoic_acid_cecum ~ Group) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj") %>% 
  add_xy_position(x = "Group") %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE)


plot_prop_final <- 
  plot_prop + 
  stat_pvalue_manual(stat.test, label = "p.adj.signif", tip.length = 0.01, hide.ns = TRUE, size = 3.5, y.position = c(1800,2000)) 


filename <- paste0("plots/Metabolites/propanoic_acid_cecum.png")
plot_prop_final 
suppressMessages(ggsave(filename = filename, width = 1800, height = 1400, units = "px", dpi = "print", device = "png", ))


# Clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```


# Butanoic acid (cecum){.tabset .tabset-fade .tabset-pills}

## Prepare dataset 
```{r prepare_butanoic, eval = TRUE}

# Load data 
metadat <- read_csv("input/input_PRIMA06_animals.csv")

# Remove samples with incomplete metadata (i.e., remove all days except D11)
dat_but <- metadat[!is.na(metadat[,"Butanoic_acid_cecum"]),]

# Turn relevant variables into factors and make sure data is in correct order 
dat_but <- dat_but %>% mutate_at("Group", factor, levels = c("Control", "Low", "Medium", "High")) %>% mutate_at("Day", factor, levels = c("D11")) %>% mutate_at("Animal", factor) 

# Identify outliers 
out <- boxplot.stats(dat_but$Butanoic_acid_cecum)$out
out_ind <- which(dat_but$Butanoic_acid_cecum %in% c(out))
out_ind

dat_but[out_ind, ] # no outliers

# Test normality 
shapiro_test(dat_but, Butanoic_acid_cecum) # Normally distributed

ggqqplot(dat_but,x = "Butanoic_acid_cecum")  

# Levene's test (test for homogeneity)
dat_but %>% 
  levene_test(Butanoic_acid_cecum~Group) # Assumption is met

```

## Visualization + summary statistics 
```{r visualize_butanoic, eval = TRUE}

# Get summary statistics 
dat_but %>% 
  group_by(Group) %>% 
  get_summary_stats(Butanoic_acid_cecum, type = "mean_sd") 

# Visualize data 
plot_but <-ggboxplot(dat_but, 
                     x = "Group", 
                     y = "Butanoic_acid_cecum",
                     fill = "Group", 
                     palette = "Pastel1", 
                     outlier.shape = NA) + 
  geom_point(size = 0.5, color = "black") +
  ylab("ug/g cecum")+
  xlab(NULL)+
  ggtitle("Butanoic acid in cecum")+
  scale_y_continuous(limits=c(700, 3500), breaks = scales::pretty_breaks(n=10))+
  theme_classic()+
  theme(legend.position="none")+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

```

## Statistical testing 

```{r test_butanoic, eval = TRUE}

# ANOVA 
anova <- aov(Butanoic_acid_cecum ~ Group, data = dat_but)
summary(anova) # not significant 


filename <- paste0("plots/Metabolites/butanoic_acid_cecum.png")
plot_but 
suppressMessages(ggsave(filename = filename, width = 1800, height = 1400, units = "px", dpi = "print", device = "png", ))


# Clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```

# Acetic acid (change){.tabset .tabset-fade .tabset-pills}

## Prepare dataset 
```{r prepare_acetic_feces, eval = TRUE}

# Load data 
metadat <- read_csv("input/input_PRIMA06_animals.csv")

# Remove samples with incomplete metadata (i.e., remove all days except D11)
dat <- metadat[!is.na(metadat[,"Acetic_acid_change"]),]

# Turn relevant variables into factors and make sure data is in correct order 
dat <- dat %>% mutate_at("Group", factor, levels = c("Control", "Low", "Medium", "High")) %>% mutate_at("Day", factor, levels = c("D07")) %>% mutate_at("Animal", factor) 

# Identify outliers 
out <- boxplot.stats(dat$Acetic_acid_change)$out
out_ind <- which(dat$Acetic_acid_change %in% c(out))
out_ind

dat[out_ind, ] # No outliers 

# Test normality 
shapiro_test(dat, Acetic_acid_change) # Normally distributed
ggqqplot(dat,x = "Acetic_acid_change") 

# Levene's test (test for homogeneity) 
dat %>% 
  levene_test(Acetic_acid_change~Group) # Assumption is met 
```

## Visualization + summary statistics 
```{r visualize_acetic_feces, eval = TRUE}

# Get summary statistics 
dat %>% 
  group_by(Group) %>% 
  get_summary_stats(Acetic_acid_change, type = "mean_sd") 

# Visualize data (new version)
plot <-ggboxplot(dat, 
                 x = "Group", 
                 y = "Acetic_acid_change",
              fill = "Group", 
              palette = "Pastel1", 
              outlier.shape = NA) + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(-1300, 3200), breaks = scales::pretty_breaks(n=10))+
  ylab("Change (ug/g feces)")+
  xlab(NULL)+
  ggtitle("Fecal acetic acid: Change between day 3 and 7")+
  theme_classic()+
  theme(legend.position="none")+ 
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))


```

## Statistical testing 

```{r test_acetic_feces, eval = TRUE}

# ANOVA
anova <- aov(Acetic_acid_change ~ Group, data = dat)
summary(anova) # Not significant 

filename <- paste0("plots/Metabolites/acetic_acid_change.png")
plot
suppressMessages(ggsave(filename = filename, width = 1800, height = 1400, units = "px", dpi = "print", device = "png", ))


# Clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```

# Propanoic acid (change){.tabset .tabset-fade .tabset-pills}

## Prepare dataset 
```{r prepare_propanoic_feces, eval = TRUE}

# Load data 
metadat <- read_csv("input/input_PRIMA06_animals.csv")

# Remove samples with incomplete metadata (i.e., remove all days except D11)
dat <- metadat[!is.na(metadat[,"Propanoic_acid_change"]),]

# Turn relevant variables into factors and make sure data is in correct order 
dat <- dat %>% mutate_at("Group", factor, levels = c("Control", "Low", "Medium", "High")) %>% mutate_at("Day", factor, levels = c("D07")) %>% mutate_at("Animal", factor) 

# Identify outliers 
out <- boxplot.stats(dat$Propanoic_acid_change)$out
out_ind <- which(dat$Propanoic_acid_change %in% c(out))
out_ind

dat[out_ind, ] # No outliers 

# Test normality 
shapiro_test(dat, Propanoic_acid_change) # Normally distributed
ggqqplot(dat,x = "Propanoic_acid_change") 

# Levene's test (test for homogeneity) 
dat %>% 
  levene_test(Propanoic_acid_change~Group) # Assumption is met 

```

## Visualization + summary statistics 
```{r visualize_propanoic_feces, eval = TRUE}

# Get summary statistics 
dat %>% 
  group_by(Group) %>% 
  get_summary_stats(Propanoic_acid_change, type = "mean_sd") 

# Visualize data (new version)
plot <-ggboxplot(dat, 
                 x = "Group", 
                 y = "Propanoic_acid_change",
              fill = "Group", 
              palette = "Pastel1", 
              outlier.shape = NA) + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(-200, 500), breaks = scales::pretty_breaks(n=10))+
  ylab("Change (ug/g feces)")+
  xlab(NULL)+
  ggtitle("Fecal propanoic acid: Change between day 3 and 7")+
  theme_classic()+
  theme(legend.position="none")+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

```

## Statistical testing 

```{r test_propanoic_feces, eval = TRUE}

# ANOVA 
anova <- aov(Propanoic_acid_change ~ Group, data = dat)
summary(anova) # Not significant


filename <- paste0("plots/Metabolites/propanoic_acid_change.png")
plot 
suppressMessages(ggsave(filename = filename, width = 1800, height = 1400, units = "px", dpi = "print", device = "png", ))


# Clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```

# Butanoic acid (change){.tabset .tabset-fade .tabset-pills}

## Prepare dataset 
```{r prepare_butanoic_feces, eval = TRUE}

# Load data 
metadat <- read_csv("input/input_PRIMA06_animals.csv")

# Remove samples with incomplete metadata (i.e., remove all days except D11)
dat <- metadat[!is.na(metadat[,"Butanoic_acid_change"]),]

# Turn relevant variables into factors and make sure data is in correct order 
dat <- dat %>% mutate_at("Group", factor, levels = c("Control", "Low", "Medium", "High")) %>% mutate_at("Day", factor, levels = c("D07")) %>% mutate_at("Animal", factor) 

# Identify outliers 
out <- boxplot.stats(dat$Butanoic_acid_change)$out
out_ind <- which(dat$Butanoic_acid_change %in% c(out))
out_ind

dat[out_ind, ] 

# Test normality 
shapiro_test(dat, Butanoic_acid_change) # Normally distributed
ggqqplot(dat,x = "Butanoic_acid_change") 

# Levene's test (test for homogeneity) 
dat %>% 
  levene_test(Butanoic_acid_change~Group) # Assumption is met 

```

## Visualization + summary statistics 
```{r visualize_butanoic_feces, eval = TRUE}

# Get summary statistics 
dat %>% 
  group_by(Group) %>% 
  get_summary_stats(Butanoic_acid_change, type = "mean_sd") 

# Visualize data (new version)
plot <-ggboxplot(dat, 
                 x = "Group", 
                 y = "Butanoic_acid_change",
              fill = "Group", 
              palette = "Pastel1", 
              outlier.shape = NA) + 
  geom_point(size = 0.5, color = "black") +
  scale_y_continuous(limits=c(-900, 1200), breaks = scales::pretty_breaks(n=10))+
  ylab("Change (ug/g feces)")+
  xlab(NULL)+
  ggtitle("Fecal butanoic acid: Change between day 3 and 7")+
  theme_classic()+
  theme(legend.position="none")+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

```

## Statistical testing 

```{r test_butanoic_feces, eval = TRUE}

# ANOVA  
anova <- aov(Butanoic_acid_change ~ Group, data = dat)
summary(anova) # Not significant 

filename <- paste0("plots/Metabolites/butanoic_acid_change.png")
plot 
suppressMessages(ggsave(filename = filename, width = 1800, height = 1400, units = "px", dpi = "print", device = "png", ))


# Clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```

# Isobutyric acid (cecum) {.tabset .tabset-fade .tabset-pills}
## Prepare dataset 
```{r prepare_isobutyric, eval = TRUE}

# Load data 
metadat <- read_csv("input/input_PRIMA06_animals.csv")

# Remove samples with incomplete metadata (i.e., remove all days except day 11)
dat_isoB <- metadat[!is.na(metadat[,"Isobutyric_acid_cecum"]),]

# Turn relevant variables into factors and make sure data is in correct order
dat_isoB <- dat_isoB %>% mutate_at("Group", factor, levels = c("Control", "Low", "Medium", "High")) %>% mutate_at("Day", factor, levels = c("D11")) %>% mutate_at("Animal", factor) 

# Identify outliers 
out <- boxplot.stats(dat_isoB$Isobutyric_acid_cecum)$out
out_ind <- which(dat_isoB$Isobutyric_acid_cecum %in% c(out))
out_ind

dat_isoB[out_ind, ] 

# Test normality 
shapiro_test(dat_isoB, Isobutyric_acid_cecum) # Normally distributed

ggqqplot(dat_isoB,x = "Isobutyric_acid_cecum")  

# Levene's test (test for homogeneity)
dat_isoB %>% 
  levene_test(Isobutyric_acid_cecum~Group) # Assumption is met

```

## Visualization + summary statistics 
```{r visualize_isobutyric, eval = TRUE}

# Get summary statistics 
dat_isoB %>% 
  group_by(Group) %>% 
  get_summary_stats(Isobutyric_acid_cecum, type = "common") 

# Visualize data 
plot_isoB <-ggboxplot(dat_isoB, 
                     x = "Group", 
                     y = "Isobutyric_acid_cecum",
                     fill = "Group", 
                     palette = "Pastel1", 
                     outlier.shape = NA) + 
  geom_point(size = 0.5, color = "black") +
  ylab("ug/g cecum")+
  xlab(NULL)+
  ggtitle("Isobutyric acid in cecum")+
  #scale_y_continuous(limits=c(700, 3500), breaks = scales::pretty_breaks(n=10))+
  theme_classic()+
  theme(legend.position="none")+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

```

## Statistical testing 

```{r isobutyric_stat, eval = TRUE}

# ANOVA 
anova <- aov(Isobutyric_acid_cecum ~ Group, data = dat_isoB)
summary(anova) # not significant 

filename <- paste0("plots/Metabolites/Isobutyric_acid_cecum.png")
plot_isoB
suppressMessages(ggsave(filename = filename, device = "png", ))

# Clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```

# Isovaleric acid (cecum) {.tabset .tabset-fade .tabset-pills}
## Prepare dataset 
```{r prepare_isovaleric, eval = TRUE}

# Load data 
metadat <- read_csv("input/input_PRIMA06_animals.csv")

# Remove samples with incomplete metadata (i.e., remove all days except day 11)
dat_isoV <- metadat[!is.na(metadat[,"Isovaleric_acid_cecum"]),]

# Turn relevant variables into factors and make sure data is in correct order
dat_isoV <- dat_isoV %>% mutate_at("Group", factor, levels = c("Control", "Low", "Medium", "High")) %>% mutate_at("Day", factor, levels = c("D11")) %>% mutate_at("Animal", factor) 

# Identify outliers 
out <- boxplot.stats(dat_isoV$Isovaleric_acid_cecum)$out
out_ind <- which(dat_isoV$Isovaleric_acid_cecum %in% c(out))
out_ind

dat_isoV[out_ind, ] 

# Test normality 
shapiro_test(dat_isoV, Isovaleric_acid_cecum) # NOT normally distributed

ggqqplot(dat_isoV,x = "Isovaleric_acid_cecum")  

# Levene's test (test for homogeneity)
dat_isoV %>% 
  levene_test(Isovaleric_acid_cecum~Group) # Assumption is met

```

### Visualization + summary statistics 
```{r visualize_isovaleric, eval = TRUE}

# Get summary statistics 
dat_isoV %>% 
  group_by(Group) %>% 
  get_summary_stats(Isovaleric_acid_cecum, type = "common") 

# Visualize data 
plot_isoV <-ggboxplot(dat_isoV, 
                     x = "Group", 
                     y = "Isovaleric_acid_cecum",
                     fill = "Group", 
                     palette = "Pastel1", 
                     outlier.shape = NA) + 
  geom_point(size = 0.5, color = "black") +
  ylab("ug/g cecum")+
  xlab(NULL)+
  ggtitle("Isovaleric acid in cecum")+
  #scale_y_continuous(limits=c(700, 3500), breaks = scales::pretty_breaks(n=10))+
  theme_classic()+
  theme(legend.position="none")+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

```

### Statistical testing 

```{r isovaleric_stat, eval = TRUE}

# Kruskal wallis 
kruskal <- kruskal.test(Isovaleric_acid_cecum ~ Group, data = dat_isoV) 

kruskal # Significant 

# Between group comparison: Dunn's test
stat.test <- dat_isoV %>%
  dunn_test(Isovaleric_acid_cecum ~ Group) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj") %>% 
  add_xy_position(x = "Group") %>%
  p_format("p.adj", accuracy = 0.0001, trailing.zero = TRUE, new.col = TRUE)

stat.test # No comparisons significant after correction 

filename <- paste0("plots/Metabolites/isovaleric_acid_cecum.png")
plot_isoV
suppressMessages(ggsave(filename = filename, device = "png", ))

# Clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```

# Isobutyric acid (change){.tabset .tabset-fade .tabset-pills}
## Prepare dataset 
```{r prepare_isob_feces, eval = TRUE}

# Load data 
metadat <- read_csv("input/input_PRIMA06_animals.csv")

# Remove samples with incomplete metadata (i.e., remove all days except D07)
dat <- metadat[!is.na(metadat[,"Isobutyric_acid_change"]),]

# Turn relevant variables into factors and make sure data is in correct order 
dat <- dat %>% mutate_at("Group", factor, levels = c("Control", "Low", "Medium", "High")) %>% mutate_at("Day", factor, levels = c("D07")) %>% mutate_at("Animal", factor) 

# Identify outliers 
out <- boxplot.stats(dat$Isobutyric_acid_change)$out
out_ind <- which(dat$Isobutyric_acid_change %in% c(out))
out_ind

dat[out_ind, ] 

# Test normality 
shapiro_test(dat, Isobutyric_acid_change) # Normally distributed
ggqqplot(dat,x = "Isobutyric_acid_change") 

# Levene's test (test for homogeneity) 
dat %>% 
  levene_test(Isobutyric_acid_change~Group) # Assumption is met 

```

## Visualization + summary statistics 
```{r visualize_isob_feces, eval = TRUE}

# Get summary statistics 
dat %>% 
  group_by(Group) %>% 
  get_summary_stats(Isobutyric_acid_change, type = "mean_sd") 

# Visualize data (new version)
plot <-ggboxplot(dat, 
                 x = "Group", 
                 y = "Isobutyric_acid_change",
              fill = "Group", 
              palette = "Pastel1", 
              outlier.shape = NA) + 
  geom_point(size = 0.5, color = "black") +
  #scale_y_continuous(limits=c(-900, 1200), breaks = scales::pretty_breaks(n=10))+
  ylab("Change (ug/g feces)")+
  xlab(NULL)+
  ggtitle("Fecal isobutyric acid: Change between day 3 and 7")+
  theme_classic()+
  theme(legend.position="none")+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

```

## Statistical testing 

```{r test_isob_feces, eval = TRUE}

# ANOVA  
anova <- aov(Isobutyric_acid_change ~ Group, data = dat)
summary(anova) # Not significant 

filename <- paste0("plots/Metabolites/Isobutyric_acid_change.png")
plot 
suppressMessages(ggsave(filename = filename, width = 1800, height = 1400, units = "px", dpi = "print", device = "png", ))


# Clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```


# Isovaleric acid (change){.tabset .tabset-fade .tabset-pills}

## Prepare dataset 
```{r prepare_isov_feces, eval = TRUE}

# Load data 
metadat <- read_csv("input/input_PRIMA06_animals.csv")

# Remove samples with incomplete metadata (i.e., remove all days except D07)
dat <- metadat[!is.na(metadat[,"Isovaleric_acid_change"]),]

# Turn relevant variables into factors and make sure data is in correct order 
dat <- dat %>% mutate_at("Group", factor, levels = c("Control", "Low", "Medium", "High")) %>% mutate_at("Day", factor, levels = c("D07")) %>% mutate_at("Animal", factor) 

# Identify outliers 
out <- boxplot.stats(dat$Isovaleric_acid_change)$out
out_ind <- which(dat$Isovaleric_acid_change %in% c(out))
out_ind

dat[out_ind, ] 

# Test normality 
shapiro_test(dat, Isovaleric_acid_change) # Normally distributed
ggqqplot(dat,x = "Isovaleric_acid_change") 

# Levene's test (test for homogeneity) 
dat %>% 
  levene_test(Isovaleric_acid_change~Group) # Assumption is met 

```

## Visualization + summary statistics 
```{r visualize_isov_feces, eval = TRUE}

# Get summary statistics 
dat %>% 
  group_by(Group) %>% 
  get_summary_stats(Isovaleric_acid_change, type = "mean_sd") 

# Visualize data (new version)
plot <-ggboxplot(dat, 
                 x = "Group", 
                 y = "Isovaleric_acid_change",
              fill = "Group", 
              palette = "Pastel1", 
              outlier.shape = NA) + 
  geom_point(size = 0.5, color = "black") +
  #scale_y_continuous(limits=c(-900, 1200), breaks = scales::pretty_breaks(n=10))+
  ylab("Change (ug/g feces)")+
  xlab(NULL)+
  ggtitle("Fecal isovaleric acid: Change between day 3 and 7")+
  theme_classic()+
  theme(legend.position="none")+
  stat_summary(fun = "mean", geom = "point", shape = 4, size = 2, position = position_dodge(.75))

```

## Statistical testing 

```{r test_isov_feces, eval = TRUE}

# ANOVA  
anova <- aov(Isovaleric_acid_change ~ Group, data = dat)
summary(anova) # Not significant 

filename <- paste0("plots/Metabolites/Isovaleric_acid_change.png")
plot 
suppressMessages(ggsave(filename = filename, width = 1800, height = 1400, units = "px", dpi = "print", device = "png", ))


# Clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```
# Heat map SCFAs vs TT
```{r heatmap, eval = TRUE}

# Load data 
dat <- read_csv("input/input_heatmap.csv")
dat_1 <- dat

colnames(dat_1)[colnames(dat_1) == "TT_D10"] = "Transit time (Day 10)"
colnames(dat_1)[colnames(dat_1) == "Acetic_acid_cecum"] = "Acetic acid (cecum)"
colnames(dat_1)[colnames(dat_1) == "Propanoic_acid_cecum"] = "Propanoic acid (cecum)"
colnames(dat_1)[colnames(dat_1) == "Butanoic_acid_cecum"] = "Butanoic acid (cecum)"

# create vector with the relevant variables (can also be subset by column indeces)
CON.VARS <- c("Transit time (Day 10)", "Acetic acid (cecum)", "Propanoic acid (cecum)", "Butanoic acid (cecum)")

# Plot all variables against each other
pairs(dat_1[,CON.VARS], pch = 19,  cex = 0.5,
      lower.panel=NULL)

# Run Pearson test
(corrmat <- cor(dat_1[,CON.VARS], method = "spearman", use = "complete.obs"))

# Create heatmap
corrmat_rounded <- round(corrmat, 2)

melted_corrmat_rounded <- tibble(Var1 = rep(row.names(corrmat_rounded), length(row.names(corrmat_rounded))),
                                 Var2 = rep(row.names(corrmat_rounded), each = length(row.names(corrmat_rounded))),
                                 dist = as.numeric(matrix(corrmat_rounded)))

filtered_data <- melted_corrmat_rounded %>%mutate_if(is.factor, as.numeric) %>% 
  filter(Var2 > Var1)


plot <- ggplot(filtered_data, aes(x = Var1, y = Var2, fill = dist)) + 
  labs(title = "Correlation between transit time and SCFAs") + 
  geom_tile(color = "white") + 
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, limit = c(-1,1), 
                      space = "Lab", name = "Correlation coefficient") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, size = 12, hjust = 1, vjust = 1), 
        axis.text.y = element_text(size = 12, hjust = 1, vjust = 0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.ticks = element_blank()) + 
  geom_text(aes(x = Var1, y = Var2, label = dist), color = "black", size = 4) + 
  coord_fixed()

# Test correlations 
with(dat, cor.test(TT_D10,Propanoic_acid_cecum, method = "spearman", use = "complete.obs"))

with(dat, cor.test(TT_D10,Acetic_acid_cecum, method = "spearman", use = "complete.obs"))

with(dat, cor.test(TT_D10,Butanoic_acid_cecum, method = "spearman", use = "complete.obs"))

# Adjust p-valyes 
p.values <- c(.0001969,.635,.03945)
p.adjust(p.values, method = "fdr")

# Save plot 
filename <- paste0("plots/Metabolites/heatmap_TT_scfas.png")
plot
suppressMessages(ggsave(filename = filename, width = 1800, height = 1400, units = "px", dpi = "print", device = "png", ))

# Clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```